### 2022-01-30 bps_conference.Rmd
data file completed

---
title: "Testing the precision of Bangla Communication Scale: Item Response Theory Guided Analysis"
output: powerpoint_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r pacman, include=FALSE}
#You need to run this chunk 1X in your system 
#install.packages("pacman")#Run this if you don't have pacman installed.
pacman::p_load(MOTE, tidyverse, psych, lavaan,kableExtra,gt,gtsummary, mirt,likert,kutils,semPlot,semTable,semTools,ggcorrplot,dlookr,qgraph,paran,EFA.MRFA,VIM,DiagrammeR,DiagrammeRsvg,ggplot2,cowplot,rsvg,questionr,magick, simsem,readxl, stringr,RColorBrewer,WrightMap, foreign,DT,SimDesign)
library(gtExtras)
library(tabledown)

#if you dont have LaTeX installed run the following lines
#install.packages("tinytex")
#tinytex::install_tinytex()
```

## Slide with Bullets

- Introduction
- Method
- Results
- Discussion

## Introduction

## Method

## Demographics
```{r demo}
knitr::include_graphics("Figures/demographics.png", dpi = 600) 
```

## Response Distribution

```{r}
knitr::include_graphics("Figures/communication.png", dpi = 600)

```

## Item Response Theory
-R version 4.1.2 (2021-11-01) with "mirt" package  
-Estimation Method: Marginal Maximum Likelihood  method with the MHRM algorithm.
_ Model : "Graded Response Model"
- Monty Carlo Study Showed for 23 items root mean squared error become stabilized for a sample range 200-250. Our sample is within that range.

## Option Charecteristics Curve
```{r}
knitr::include_graphics("Figures/problem.icc.png", dpi = 600)

```

## Option Charecteristics Curve (Cnt.)
```{r}
knitr::include_graphics("Figures/good.icc.png", dpi = 600)

```

## Item Plots

```{r ip1}
knitr::include_graphics("Figures/ICC1.png", dpi = 600) 
```

## Item Plots (Cont)

```{r ip2}
knitr::include_graphics("Figures/ICC2.png", dpi = 600) 
```

## Item Plots (Cont)

```{r ip3}
knitr::include_graphics("Figures/ICC3.png", dpi = 600) 
```

## Item Plots (Cont)

```{r ip4}
knitr::include_graphics("Figures/ICC4.png", dpi = 600) 
```

## IRT Parameters
```{r irtparadata, include=F}
param <- readr::read_csv("table_raw/IRT_parameters.csv")

```

## Person Fit

```{r personfit}
knitr::include_graphics("Figures/person.png", dpi = 600)

```

```{r para, results='asis'}
param %>% 
  flextable::flextable() 
```

## Test Information Plot

```{r TIC, results='asis'}
knitr::include_graphics("Figures/TIC.png", dpi = 600)

```

## Conditional Reliability
```{r conrel}
knitr::include_graphics("Figures/conrel.png", dpi = 600)

```