### 2022-02-02 manuscript.Rmd
Introduction part updated

---
title             : "Psychometric Evaluation of the Bangla-translated Rotter Internal-External Scale through Classical Test Theory and Item Response Theory"
shorttitle        : "Bangla Communication Scale"

author: 
  - name          : "Nusrat Jahan"
    affiliation   : "1"
    corresponding : no    # Define only one corresponding author
    role:         
      - Conceptualization
      - Data Curation
      - Writing - Original Draft Preparation
  - name          : "Mushfiqul Anwar Siraji"
    affiliation   : "2"
    corresponding : no   # Define only one corresponding author
    address       : "Monash University Malaysia,Jalan Lagoon Selatan, 47500 Bandar Sunway,Selangor Darul Ehsan, Malaysia."
    email         : "mushfiqul.anwarsiraji@moansh.edu"
    role:
      - Conceptualization
      - Project Management
      - Formal Analysis
      - Data Visualization
      - Writing - Original Draft Preparation
      - Writing - Review & Editing

  - name          : "Zinnatul Borak"
    affiliation   : "1"
    corresponding : yes   # Define only one corresponding author
    address       : "Department of Educational and Counselling Psychology,University of Dhaka,Dhaka 1000"
    email         : "mushfiqul.anwarsiraji@moansh.edu"
    role:
      - Conceptualization
      - Writing - Review & Editing

affiliation:
  - id            : "1"
    institution   : "Department of Educational and Counselling Psychology,University of Dhaka"
  - id            : "2"
    institution   : "Monash University, Department of Psychology, Jeffrey Cheah School of Medicine and Health Sciences, Malaysia"

authornote: |
  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

  Enter author note here.

abstract: |
  One or two sentences providing a **basic introduction** to the field,  comprehensible to a scientist in any discipline.

  Two to three sentences of **more detailed background**, comprehensible  to scientists in related disciplines.

  One sentence clearly stating the **general problem** being addressed by  this particular study.

  One sentence summarizing the main result (with the words "**here we show**" or their equivalent).

  Two or three sentences explaining what the **main result** reveals in direct comparison to what was thought to be the case previously, or how the  main result adds to previous knowledge.

  One or two sentences to put the results into a more **general context**.

  Two or three sentences to provide a **broader perspective**, readily comprehensible to a scientist in any discipline.

  <!-- https://tinyurl.com/ybremelq -->

keywords          : "keywords"
wordcount         : "X"

bibliography      : ["r-references.bib", "references.bib"]

floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
---

```{r setup, warning=FALSE, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = F, message = F, dpi=1200)

set.seed(123)
par(family = "Arial")
```

```{r pacman, eval=FALSE, include=FALSE}
#You need to run this chunk 1X in your system 
#install.packages("pacman")#Run this if you don't have pacman installed.
pacman::p_load(MOTE, tidyverse, psych, lavaan,kableExtra,gt,gtsummary, mirt,likert,kutils,semPlot,semTable,semTools,ggcorrplot,dlookr,qgraph,paran,EFA.MRFA,VIM,DiagrammeR,DiagrammeRsvg,ggplot2,cowplot,rsvg,questionr,magick, simsem,readxl, stringr,RColorBrewer,WrightMap, foreign,DT,SimDesign)
library(papaja)
pacman::p_install_gh("jthomasmock/gtExtras")
pacman::p_install_gh("crsh/papaja")
pacman::p_install_gh("masiraji/tabledown")
pacman::p_install_gh("crsh/citr")

papaja::r_refs("references.bib")
#if you dont have LaTeX installed run the following lines
#install.packages("tinytex")
#tinytex::install_tinytex()
renv::snapshot()
```

```{r library,include=FALSE}
#This chunk holds code for calling the required packages

library(papaja) # devtools::install_github("crsh/papaja")
library(lavaan)
library(semPlot) #devtools::install_github('SachaEpskamp/semPlot',  dependencies = T)
library(semTools)
library(MOTE)
library(car)#reversecoding 
library(psych)
library(dlookr)
library(plyr)
library(tidyverse)
library(qgraph)
library(kableExtra)
library(paran) # Parallel analysis
library(EFA.MRFA) # Hull method
library(DiagrammeR) #devtools::install_github('rich-iannone/DiagrammeR')
library(DiagrammeRsvg) #For DiagrammeR
library(rsvg) #For DiagrammeR
library(ggcorrplot)
library(semTable)
library(magick)
library(mirt)
library(gtsummary)
library(gt)
library(gtExtras)
library(likert)
library(VIM) #Missing data
library(kutils)
library(simsem)
library(tabledown)
library(readxl)
library(koRpus)# stable release
library(koRpus.lang.en)
library(reshape)
library(ggsci)
renv::snapshot()
r_refs("references.bib")
```

---
nocite: |
  `r papaja::cite_r("references.bib")`

---

Communication is a complex behaviour of exchanging information among individuals [@tanner2006advanced]. Communication plays a central role among adolescents in developing self-identity, social relationships and creates the foundation of collective social activity [@haslett1989children; @conti2008emotional; @spencer2013language]. Inadequate communication skill may cause poor peer relationship resulting long-term socio-emotional difficulties including social anxiety, stress, low self-esteem and poor academic performance [@brinton2004social; @reed2020relative].

Often adults picture adolescents having inadequate and inept communication skills [@thurlow2003teenagers;@stern2005self]. Media representation of adolescents often includes "storm-and stress", self absorbed and disengaged type behaviours [@porteous1980adolescents;@stern2005self]. As such adolescents are often labelled as "lazy" and "disrespectful" by the adults [@public1999kids]. On the contrary adolescents are highly engaged in work, community services and extracurricular activities and also more aspiring to earn an college degrees [@debard2004millennials; @schneider1999ambitious]. Such a discrepancy between the reality of adolescent's image and adult's perception of the adolescents might be attributed to the mismatch of communication skills. The communication pattern of adolescents might not necessarily same as the adults. In addition to face to face communication, adolescents are vastly exposed to different virtual communication platforms. This may cause them to face more complex social challenges than the adults [@thurlow2003teenagers]. "Communication capital" expresses the potential of civic-engagement that incorporates developing social relationships and influences collective social activity. The more communication capital an individual has the easier the instances of civic engagement become.

Understanding the adolescents' communication skill  vital as it is considered as the "key skill" in the education [@thurlow2001talkin] and employment market [@olszewski2017training]. In the western society, adolescents are now facing high unemployment [@lindsay2014employers]. Lack of adequate communication skill is one of the root causes of this high unemployment [@lindsay2014employers]. Similarly, lack of proper communication skill often promotes the propensity of anti-social behaviours and risk of exclusion from schools [@clegg2009language; @conti2004social].

To promote better understanding of subject contents assessing the communication skill among adolescents is highly required. For this purpose "Communication Skill" sub-skill set [@barkman2002four]  was developed in 2002 as a part of The  National  On-line  Youth  Life  Skills  Evaluation System [@mincemoyer2005measuring] and since then it has been extensively used [@fitzpatrick2005life].

This study has two objectives: (a) To explore and validate the latent construct of "Bangla Communication Scale" using (b) To gather concurrent validity evidence (c) To increase the precision of the scale usining Item Response Theory guided analysis
## Methods

## Participants

```{r data, include=F}
# data <- readxl::read_excel("Processeddata/communication_data_n_582.xlsx")
# readr::write_rds(data, "Processeddata/communication_data_n_582.rds")
data <- readr::read_rds("Processeddata/communication_data_n_582.rds")

efa.data.full <- subset(data, Sample == "EFA")
cfa.data.full <- subset(data, Sample == "CFA"| Sample=="Validity")
validity.data.full <- subset(data, Sample == "Validity")

sem.data <- data[,c(2,6:66)]
efa.data <- subset(sem.data , Sample == "EFA")
efa.data <- efa.data[,c(2:24)]
efa.data <- na.omit(efa.data)

cfa.data<- subset(sem.data , Sample == "CFA"| Sample=="Validity")
cfa.data <- cfa.data[,c(2:24)]
cfa.data <- na.omit(cfa.data)

validity.data <- subset(sem.data , Sample == "Validity")
validity.data <-na.omit(validity.data)

irt.data <- sem.data[,2:24]
```

```{r sescrptives, include=FALSE}
desc <- efa.data.full[,c(3,4,5)]
## Recoding desc$Sex
desc$Gender <- fct_recode(desc$Gender,
  "Boy" = "male",
  "Girl" = "female"
)
colnames(desc) <- c("Gender", "Age", "Socio Economic Status")

des_tab <- desc %>%
  tbl_summary(
    by = Gender,
    statistic = list(all_continuous() ~ "{mean} ({sd})",
                     all_categorical() ~ "{n} ({p}%)"),
    type = list(Age ~ 'continuous'),
    digits = all_continuous() ~ 2,

    missing = "no"
    ) %>% add_overall() %>% bold_labels() %>% 
modify_header(label ~ "**Variable**")  %>%
  modify_caption("Demographic Characteristics")

as_tibble(des_tab) -> desc_tibble 
as_kable_extra(des_tab, format = "latex",booktabs = T) -> desc_kable #save it as a knitr::kable

des.2 <- desc %>% 
  psych::describeBy("Gender")
female <- as.tibble(des.2$Girl)
male <- as.tibble(des.2$Boy)

```

A cross-sectional survey was used to collect data from a large sample of students of grade 8-12 (n = `r nrow(efa.data)`) from 8 schools following convenience sampling method. T. For estimating the sample size for the CFA we followed the N:q rule [@comreyFirstCourseFactor1992; @schonbrodtWhatSampleSize2013]  which required 10 participants per parameters for trustworthiness of the result. Our sample size exceeds the requirement. Among `r nrow(efa.data)` participants, `r female[1,2]` were female aged between `r female[2,8]` to `r female[2,9]` years (`r round(female[2,3],2)`±`r round(female[2,4],2)`). `r male[1,2]`  were male with an age range between `r male[2,8]` participants `r male[2,9]` years (`r round(male[2,3],2)`±`r round(male[2,4],2)`). `r desc_tibble [4,2]` participants belonged to middle socio-economic status. `r desc_tibble [3,2]` and `r desc_tibble [5,2]` participants belonged to lower and upper socio-economic status-respectively.

## Procedure
Prior to data collection necessary authorization from school's authority and assent from the participating were obtained. Data collection was commenced between November 2021 to January 2022. The data collection took place in the classroom where students were at first briefed about 'communication skill'. Next. they filled up their soci-demographics information and responded to our Bangla Communication Scale. All personal information (name, school,class) was codified and encrypted, producing a anonymous database.

## Materials

### Communication Scale

### Bangla Communication Scale
We translated the "Communication Scale" into Bangla Language following  International Test Commission (ITC) guidelines [@bartramITCGuidelinesTranslating2018] to translate and adapt Communication scale. Two bilingual researchers (PhD in Psychology) natives in Bangla translated the original English version to Bangla. Two translated versions were then judged and synthesized by the authors. Subsequently, two bilingual researchers (One PhD, one MS in Psychology) back-translated the Bangla scale into English with no knowledge of the original work. The authors synthesized the two back-translations and compared it with the original scale and made necessary amendments

```{r, gtfig, include =F}

gt.data <- na.omit(efa.data)

names(gt.data)[names(gt.data) == "RCS02"] <- "CS02"
names(gt.data)[names(gt.data) == "RCS05"] <- "CS05"

#Creating long data
gt.data.long <- as.data.frame(gather(gt.data, Items, value))
gt.data.long$value <- as.numeric(as.character(gt.data.long$value))

gt.data.tab <- gt.data.long %>% 
  group_by(Items) %>% 
  # calculate summary stats & create data for the histogram and density plot
  dplyr::summarise(
    nr = n(),
    mean = mean(value, na.rm = TRUE),
    med = median(value, na.rm = TRUE),
    sd = sd(value, na.rm = TRUE),
    hist_data = list(value),
    dens_data = list(value),
    .groups = "drop"
  ) %>% 
  gt()

gt.likert <-  as.data.frame(gt.data)

#gt.likert <-  mutate(gt.likert, across(starts_with("ASQ"), ~unname(recod[.])))

gt.likert.Factor = as.data.frame(lapply(gt.likert,factor,
                                     ordered = T))

#get the items name
gt.items <- names(gt.likert.Factor) 
#Calculate percentage
gt.percentage <- kutils::likert(gt.likert.Factor, vlist = gt.items )

gt.percentage <- gt.percentage$table %>% 
  as.data.frame(.)

#data wrangling
labels <- c("never", "rarely","sometimes", "often", "always","Total")
as.data.frame(labels)

full.percentage <- cbind(labels,gt.percentage) #tables with labels  
full.percentage<- t(full.percentage ) #transpose
as.data.frame(full.percentage)
full.percentage1 <- full.percentage[-1,-6] #removing 1st row and total column
#full.percentage2 <- full.percentage1[, c(3, 4, 1, 2,5)]# rearranging
as.data.frame(full.percentage1)

colnames(full.percentage1) <- c("Never", "Rarely","Sometimes", "Often", "Always")

Items <- rownames(full.percentage1)
as.data.frame(Items)
full.percentage3 <- cbind(Items,full.percentage1)

full.percentage3 <- full.percentage3[order(Items),] 
full.percentage3 <- as.data.frame(full.percentage3[,-1])

full.percentage3 <- full.percentage3%>% 
  gt()

gt.table <- gt.data.tab$'_data'
gt.liket <- full.percentage3$"_data"

Table <-  cbind( gt.table, gt.liket )%>% 
  gt()%>% 
  # histogram and density plots
  gtExtras::gt_sparkline(
    hist_data,
    type = "histogram",
    line_color = "black", 
    fill_color = "#00A08799",
    bw = 1,
    same_limit = TRUE)%>%
  gtExtras::gt_sparkline(
    dens_data,
    type = "density",
    line_color = "black", 
    fill_color = "#E64B3599",
    bw = 0.75,
    same_limit = TRUE
  )%>%
  # format decimals
  fmt_number(columns = mean:sd, decimals = 1) %>%
  # header
  tab_header(
    title = md("**Communication**"),
    subtitle = md("Summary Descriptives")
  ) %>% 
  tab_footnote(
    footnote = md("**Histogram**"),
    locations = cells_column_labels(columns = hist_data)
  ) %>%
  tab_footnote(
    footnote = md("**Density**"),
    locations = cells_column_labels(columns = dens_data)
  ) %>% 
  #create groups of columns
  tab_spanner(
    label = "Items",
    columns = Items
  ) %>%
  tab_spanner(
    label = "Summary Statistics",
    columns = nr:sd
  ) %>%
  tab_spanner(
    label = "Graphics",
    columns = hist_data:dens_data
  )  %>%
  # change column names to appear in the table
  cols_label(
    Items = ("Items"),
    nr = ("n"),
    mean = ("Mean"),
    med = ("Median "),
    sd = (("SD")),
    hist_data = "Histogram",
    dens_data = "Density"
  ) %>%

  tab_spanner(
    label = "Response Pattern",
    columns = "Never":"Always"
  ) %>%
  # set alignment as per wish
  cols_align(
    align = "center",
    columns = everything()
  ) %>%
  opt_align_table_header(align = "left") %>%
  gtExtras::gt_plt_dot(
    mean,
    Items,
    palette = "ggthemes::fivethirtyeight")

gtsave(Table, "Figures/efa_communication.png",vwidth = 7000)

```

```{r ggplot2, include=F}
#This chunk holds code for creating ggplot2 based apatheme for plots.
apatheme=theme_bw()+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        #text=element_text(family = "Helvetica"),
        axis.text.x = element_text(size = 15),
        axis.text.y = element_text(size = 15),
        axis.title.x = element_text(size = 15),
        axis.title.y = element_text(size = 15),
        plot.title = element_text(size = 15),
        legend.text = element_text(size = 15),
        legend.title=element_blank(),
        axis.line.x = element_line(color='black'),
        axis.line.y = element_line(color='black'),
        panel.border = element_rect(color = "black",
                                    fill = NA,
                                    size = 1))

```

```{r EFAdata, include=FALSE}

#library(VIM)
missing <- VIM::aggr(efa.data, plot =T)

```

```{r EFAassumptions, include=FALSE}

#KMO test
KMO <- psych::KMO(efa.data)

# Test of correlation matrix
bartlet <- psych::cortest.bartlett(efa.data, n =300)

decriptives <- tabledown::des.tab(efa.data)
colnames(decriptives) = c("Items","Mean", "SD", "Skew", "Kurtosis", "Shapiro-Wilk Statistics","Item-Total Correlation")

```

```{r}
apa_table(decriptives)
```

## Item Analysis

```{r Cronbach, include=FALSE, results='asis'}
Cron.bach <- psych::alpha(efa.data,check.keys=TRUE) 
Cron.bach$total %>% 
  knitr::kable(caption = "Alpha", digits = 2) %>% 
  kableExtra::kable_styling(bootstrap_options = "striped", full_width = F)
Cron.bach$alpha.drop %>% 
  knitr::kable(caption = "Alpha value if item is dropped", digits = 2) %>% 
  kableExtra::kable_styling(bootstrap_options = "striped", full_width = F)

Cron.bach$item.stats %>% 
  knitr::kable(caption = "Corrected Item total Correlation", digits = 2) %>% 
  kableExtra::kable_styling(bootstrap_options = "striped", full_width = F)

low.r.corec <- (min(Cron.bach$item.stats$r.cor))  #minimum item total correlation
high.r.corec <- (max(Cron.bach$item.stats$r.cor))     # maximum item total correlation

```

```{r CorrMatrix, include=FALSE}
#Polychoric Correlation matrix

correlations <- psych::polychoric(efa.data)
upper <- correlations$rho[upper.tri(correlations$rho, diag = F)]

min.cor <- apa(min(abs(upper)),2,F) #minimum cor.coefficient of the matrix
max.cor <- apa(max(abs(upper)),2,F) # max. correlation coefficient of the matrix

determinets <- det(correlations$rho) 
# Calculating the percentage of correlations higher than .30

BigR = sum(correlations$rho >= abs(.30) & correlations$rho < abs(1.0), na.rm =T)/2 
totR = length(efa.data)*(length(efa.data)-1)/2
cor.per <- print (BigR/totR)*100
```

```{r corplot, include =F}
#This chunk holds code for plotting the correlation matrix

corplot <- ggcorrplot(correlations$rho, hc.order = TRUE, outline.col = "white", type = "lower", 
           ggtheme = ggplot2::theme_minimal(),tl.srt = 90,colors = c("#E64B35FF", "white", "#3C5488FF") )+
  theme( panel.grid.major = element_blank(),
         axis.text.x = element_text(size = 8),
         axis.text.y = element_text(size = 8),
         panel.border = element_rect(color = "black",
                                    fill = NA,
                                    size = 1))
#p.mat <- cor_to_p(correlations$rho, 428, method = "polychoric")
#to show pvalue in the fig add p.mat=p.mat
ggsave("Figures/corplot.png",corplot, width = 6, height = 6, dpi = 600, bg ="white")

```

(ref:figCor) Inter item polychoric correlation coefficients for the 48 items. 4.9 % inter-item correlation coefficients were higher than .30

```{r figCor, out.height="150%",out.width="100%", fig.align= 'center', fig.cap='(ref:figCor)', results='asis', fig.align='center'}

knitr::include_graphics('Figures/corplot.png', dpi =600)

```

```{r ParallelEFA, include=FALSE}
#This chunk holds code for factor identification method : Horn
# Parallel analysis
Horn <- paran(correlations$rho, iterations=500, centile=0, quietly=FALSE,
           status=TRUE, all=FALSE, cfa=FALSE, graph=TRUE,
           color=TRUE, col=c("black","red","blue"),
           lty=c(1,2,3), lwd=1, legend=TRUE, 
           width=640, height=640, grdevice=png, seed=0, mat=NA, n=NA)

Adjusted.EV <- data.frame(Horn$AdjEv)
Adjusted.EV$type = c('Adjusted Ev')
Adjusted.EV$num = c(row.names(Adjusted.EV))
Adjusted.EV$num = as.numeric(Adjusted.EV$num)
colnames(Adjusted.EV) = c('eigenvalue', 'type', 'num')

Random.EV <- data.frame(Horn$RndEv)
Random.EV$type = c('Random EV')
Random.EV$num = c(row.names(Random.EV))
Random.EV$num = as.numeric(Random.EV$num)
colnames(Random.EV) = c('eigenvalue', 'type', 'num')

Unadjusted.EV <- data.frame(Horn$Ev)
Unadjusted.EV$type = c('Unadjusted EV')
Unadjusted.EV$num = c(row.names(Unadjusted.EV))
Unadjusted.EV$num = as.numeric(Unadjusted.EV$num)
colnames(Unadjusted.EV) = c('eigenvalue', 'type', 'num')

parallel = ggplot(Adjusted.EV, aes(x=num, y=eigenvalue, shape=type, color=type)) +
  #Add lines connecting data points
  geom_line(colour ='#DC000099')+
    #geom_line(data = Random.EV) +
  #geom_line(data = Unadjusted.EV)+
    #Add the data points.
  geom_point(size=1.5,colour ='#3C5488FF')+
  #geom_point(data = Random.EV, size=3)+
  #geom_point(data = Unadjusted.EV, size=3)+
  scale_y_continuous(name="Eigen Value", limits=c(0, 4), breaks=seq(0,4,1)) +
   #Label the x-axis 'Factor Number', and ensure that it ranges from 1-max # of factors, increasing by one with each 'tick' mark.
  scale_x_continuous(name='Factor Number (Parallel Analysis)', limits=c(0, 23),breaks=seq(0,23,4))+
    #Add vertical line indicating parallel analysis suggested max # of factors to retain
  geom_hline(yintercept=1, linetype = 'dashed') + apatheme +theme(,legend.position = "None") 
#text = element_text(size = 25)

```

```{r ScreePlot, include=FALSE}

#This chunk holds code for factor identification method : Scree Plot

Scree <- scree(correlations$rho,factors=TRUE,pc=TRUE,main="(B)",
      hline=NULL,add=F)

FA.Scree <- data.frame(Scree$fv)
FA.Scree$type = c('Factor Analysis')
FA.Scree$num = c(row.names(FA.Scree))
FA.Scree$num = as.numeric(FA.Scree$num)
colnames(FA.Scree) = c('eigenvalue', 'type', 'num')

PA.Scree <- data.frame(Scree$pcv)
PA.Scree$type = c('Principal Component Analysis')
PA.Scree$num = c(row.names(PA.Scree))
PA.Scree$num = as.numeric(PA.Scree$num)
colnames(PA.Scree) = c('eigenvalue', 'type', 'num')

scree.plot <-ggplot(FA.Scree, aes(x=num, y=eigenvalue,color=tye, shape=type))+
   geom_line(colour ='#DC000099')+
  #geom_line(data = PA.Scree)+ 
  geom_point(size=1.5,colour ='#3C5488FF')+
  #geom_point(data = PA.Scree, size=3)+
  scale_y_continuous(name="Eigen Value", limits=c(0, 5), breaks=seq(0,5,1)) +
  scale_x_continuous(name='Factor Number (Scree Plot)', limits=c(0, 20),breaks=seq(0,20,4))+
  geom_hline(yintercept=1, linetype = 'dashed') + apatheme +theme(legend.position = "None") 
#text = element_text(size = 25

```

```{r HullEFA, include=FALSE}
#This chunk holds code for factor identification method : Hull

# HULL
EFA.MRFA::hullEFA(efa.data,extr = "ULS", index_hull = "CAF", display = TRUE, graph = T,
        details = TRUE)
hull <- ggplot2::last_plot()
Hull <- hull+ ggtitle(NULL)+xlab("Factor Number (Hull Method)")+ylab("CAF")+ aes(color = '#DC000099')+geom_point(color ='#3C5488FF',size =1)+
  apatheme +theme(legend.position = "None") +scale_fill_identity()

```

```{r MAPefa, include=FALSE}
#This chunk holds code for factor identification method : MAP
#MAP
map <- psych::VSS(efa.data, rotate = "promax", fm = 'pa', n.obs =300 )
map.map <- as.data.frame(map$map)
colnames(map.map) <- "MAP Statistic"
map.statistics <- map$vss.stats[,c(1,2,5,6,7,10,11)]
full.map <- cbind(map.map,map.statistics)

write.csv(full.map, "Table_raw/map_stat.csv")
```
Scree plot, map and hull method ( Figure \@ref(fig:facIdFig)) suggested a one factor solution. Horn's parallel analysis [@hornRationaleTestNumber1965] with 500 iterations  indicated a two-factor solution. However, the minimum average partial (MAP) method (Table \@ref(tab:factormp)) [@velicerDeterminingNumberComponents1976] and Hull method [@lorenzo-sevaHullMethodSelecting2011] ( Figure \@ref(fig:facIdFig)) suggested a five-factor solution. As a result, we tested both five-factor and six-factor solutions.
```{r factors, include=F}
#This chunk holds code for compiled factor identification plots

factor <- cowplot::plot_grid(parallel, scree.plot,Hull, 
                   labels = "AUTO",
                   ncol=1, 
                  align = "v", 
                  label_fontfamily = "sans",
                  label_fontface = "plain")
ggsave("Figures/factors.png",factor, width = 5, height = 7, dpi = 600)
```

(ref:facIdFig) Factor Identification Methods (A) Parallel analysis  indicated the optimal number of factors were two. (B) Scree plot suggested One facor. (C) Hull method indicated 1 factors were required to balance the model fit and number of parameters.

```{r facIdFig, fig.align= 'center', fig.cap='(ref:facIdFig)',  warning=FALSE, results= 'asis'}

knitr::include_graphics("Figures/factors.png", dpi =800)

```

```{r EFA, include=FALSE}
fa.2F.1 <- fa(r=correlations$rho, nfactors = 2, fm= "pa",rotate ="varimax",
              residuals = TRUE, SMC = TRUE)
AA <- print(fa.2F.1, cut = .3, digits = 3, sort = TRUE)

reduced.model.2F.1 <- dplyr::select(efa.data ,
                                    -c( CS21, CS11, CS23, CS09, RCS02,RCS05, CS16, CS01))

correlations.red.2F.1 <- polychoric(reduced.model.2F.1)

fa.2F.2 <- fa(r=correlations.red.2F.1$rho, nfactors = 2, fm= "pa",rotate ="varimax",
              residuals = TRUE, SMC = TRUE)

BB <- print(fa.2F.2, cut = .3, digits = 3, sort = TRUE)

reduced.model.2F.2 <- dplyr::select(efa.data ,
                                    -c( CS21, CS11, CS23, CS09, RCS02,RCS05, CS16, CS01, CS03))

correlations.red.2F.2 <- polychoric(reduced.model.2F.2)

fa.2F.3 <- fa(r=correlations.red.2F.2$rho, nfactors = 2, fm= "pa",rotate ="varimax",
              residuals = TRUE, SMC = TRUE, max.iter = 500 )

CC <- print(fa.2F.3, cut = .3, digits = 3, sort = TRUE)

var1 <- CC$Vaccounted[2,1]*100
var2 <- CC$Vaccounted[2,2]*100

omega <- psych::omega(reduced.model.2F.2,2)

```

```{r EFA2f, include=FALSE}
fa.1F.1 <- fa(r=correlations$rho, nfactors = 1, fm= "pa",
              residuals = TRUE, SMC = TRUE)
AA <- print(fa.1F.1, cut = .3, digits = 3, sort = TRUE)

reduced.model.1F.1 <- dplyr::select(efa.data ,
                                    -c( CS16, RCS02, RCS05, CS01, CS09, CS23))

correlations.red.1F.1 <- polychoric(reduced.model.1F.1)

fa.1F.2 <- fa(r=correlations.red.1F.1$rho, nfactors = 1, fm= "pa",
              residuals = TRUE, SMC = TRUE)

BB <- print(fa.1F.2, cut = .3, digits = 3, sort = TRUE)

```

```{r eval=FALSE, include=FALSE}
model.1 <- "F1 =~ CS15 + CS22+CS20+CS14+CS18+CS13+CS06+CS10
            F2 =~ CS19+CS04+CS17+CS12+CS08+CS07"
fit.1 <- cfa(model.1, data = cfa.data, ordered = names(cfa.data),estimator = "WLSMV")

## Summary of Model 
cfa.sum.1 <- summary(fit.1, fit.measures =TRUE,standardized = TRUE,rsq =TRUE)

## Selected Fit measures 
fit.measures.1 <- fitmeasures (fit.1,c("gfi", "agfi", "nfi","rfi", 
                       "cfi","tli",
                       "rmsea", "rmsea.ci.lower", "rmsea.ci.upper","srmr"))
reliability1 <- semTools::reliability(fit.1, return.total = T)

```

```{r CFAitems, include=FALSE }
#This chunk holds codes for 1st CFA

model.1 <- "Communication =~ CS15 + CS21 +CS11 + CS20+ CS14+ CS19+ CS22+ CS12+ CS06+ CS07+ CS18+ CS08+        CS04+CS17+CS03+CS13+CS10"

fit.1 <- cfa(model.1, data = cfa.data, ordered = names(cfa.data),estimator = "WLSMV")

## Summary of Model 
cfa.sum.1 <- summary(fit.1, fit.measures =TRUE,standardized = TRUE,rsq =TRUE)

## Selected Fit measures 
fit.measures.1 <- fitmeasures (fit.1,c("gfi", "agfi", "nfi","rfi", 
                       "cfi","tli",
                       "rmsea", "rmsea.ci.lower", "rmsea.ci.upper","srmr"))

reliability1 <- semTools::reliability(fit.1)

```

```{r mod1, eval=FALSE, include=FALSE}
#This chunk holds codes for modification indices for 1st CFA
modfit.Cor.one <- modindices(fit.Cor.1, sort. = TRUE) 
modfit.Cor.one[modfit.Cor.one$mi>3.84,]
```

```{r cfaplot, include =F}
#This chunk holds codes to store CFA plot

png(filename="figures/CFAplot.png", 
    type="cairo",
    units="in", 
    width=6, 
    height=6, 
    pointsize=12, 
    res=1200)
semPaths (fit.1 , 
          what= "std",
          #"hide", #(hides coefficeits)
          #whatLabels = "std",
          intercepts = F,
          style ="OpenMx",
          #residScale = 6,
          theme = "colorblind",
          nCharNodes = 0,
          reorder =T,
          rotation =2,
          layout ="tree",
          cardinal = T,
          curvePivot =T,
          sizeMan =8,#items length
          sizeMan2 = 2,#items
          sizeLat = 12,#factors
          thresholds = FALSE,
          equalizeManifests =F,
          fade = FALSE,
          edge.label.cex = .8,
          #exoCov = T,
          centerLevels = T,
          #edge.color="black",
          label.scale=T,
          label.cex=1.2, #Font of factor and item name
          residuals=FALSE,
           #exoVar=FALSE,
          #fixedStyle=6, #Style of arrow (guide item)
          #freeStyle=1#Style of arrow (other item), #XKCD = TRUE

          )

dev.off()
```

```{r Shortcut Invariance, eval=FALSE, include=FALSE}
invariance.data <- cfa.data.full[, c(3, 6:28)]
invariance.data <- na.omit(invariance.data )
library("semTools")
invariance.model <- measurementInvariance(model = model.1,
                                          data = invariance.data,
                                          group = "Gender",                                                                         estimator = "WLSMV",
                                          strict = T)
invariance.model[[1]] #configural Invariance
invariance.model[[2]] # loading, weak, metric
invariance.model[[3]] # intercept, strong, scaler
invariance.model[[4]] # residual, strict
invariance.model[[5]] # means, structural invariance
```

```{r figcfa, fig.cap = "CFA Plot.", warning=FALSE}

knitr::include_graphics("Figures/CFAplot.png", dpi =600)

```

## Measurement Invariance
```{r InvarianceAnalysis_data, include=FALSE}
#This chunk holds codes to prepare measurement invariance data.

invariance.data <- cfa.data.full[, c(3, 6:28)]
invariance.data <- na.omit(invariance.data )
mi.count <- invariance.data %>% 
  group_by(Gender) %>% 
  count() %>% 
  as.data.frame()
```

```{r Invariancedetail, include=FALSE}
#This chunk holds codes for measurement Invariance

model.1 <- "Communication =~ CS15 + CS21 +CS11 + CS20+ CS14+ CS19+ CS22+ CS12+ CS06+ CS07+ CS18+ CS08+        CS04+CS17+CS03+CS13+CS10"

# Configural invariance ####
configural <- cfa(model = model.1,
                  data = invariance.data,
                  group = "Gender",
                  estimator = "WLSMV")

summary(configural, fit.measures =TRUE,standardized = TRUE, rsq =TRUE)

fitmeasures (configural,c("gfi", "agfi", "nfi","rfi", 
                       "cfi","tli",
                       "rmsea", "rmsea.ci.lower", "rmsea.ci.upper","srmr"))

#Metric
weak <- cfa(model = model.1,
                  data = invariance.data,
                  group = "Gender",

                  estimator = "WLSMV", 
            group.equal = "loadings")

summary(weak, fit.measures =TRUE,standardized = TRUE, rsq =TRUE)

fitmeasures (weak,c("gfi", "agfi", "nfi","rfi", 
                       "cfi.scaled","tli.scaled",
                       "rmsea.scaled", "rmsea.ci.lower.scaled", "rmsea.ci.upper.scaled","srmr"))

First.comp <- compareFit (configural, weak)
summary(First.comp)

#Scaler
strong <- cfa(model = model.1,
                  data = invariance.data,
                  group = "Gender",

                  estimator = "WLSMV", 
              group.equal = c("loadings", "intercepts"))
summary(strong, fit.measures =TRUE,standardized = TRUE, rsq =TRUE)

Second.comp <- compareFit  (weak, strong)
summary(Second.comp)

#residual
strict <- cfa(model = model.1,
                  data = invariance.data,
                  group = "Gender",

                  estimator = "WLSMV",  
              group.equal = c("loadings", "intercepts", "residuals"))

summary(strict, fit.measures =TRUE,standardized = TRUE, rsq =TRUE)
residual.fitm <- as.data.frame(fitmeasures (strict,c("gfi", "agfi", "nfi","rfi", "cfi","tli","rmsea", "srmr","aic")))

Third.comp <- compareFit(strict,strong)
summary(Third.comp)
#structural
structural <-cfa(model = model.1,
                  data = invariance.data,
                  group = "Gender",

                  estimator = "WLSMV",  
                 group.equal = c("loadings", "intercepts", "residuals", "means","lv.variances","lv.covariances"))

summary(structural, fit.measures =TRUE,standardized = TRUE, rsq =TRUE)
fitmeasures (structural,c("gfi", "agfi", "nfi","rfi", "cfi","tli","rmsea", "srmr","aic"))

fourth.comp <- compareFit  (structural,strict)
summary(fourth.comp)
comfit.par <- compareFit(configural, weak, strong, strict)
summary (comfit.par)

models <-  list("Configural" = configural, 
                "Metric" = weak, 
                "Scalar" = strong, 
                "Residual" = strict)

Invariance.table <- compareLavaan(models,
              nesting = "Configural > Metric > Scalar > Residual", 
              fitmeas = c("chisq", "df",  "cfi","tli","rmsea", "rmsea.ci.lower", "rmsea.ci.upper", "srmr" ),
              scaled = F,
              chidif = T, digits = 2)

colnames(Invariance.table) <- c("Chi-Square", "df", "CFI", "TLI", "RMSEA", "RMSEA 90% Lower CI", "RMSEA 90% Upper", "Chi-Square Difference", "df difference*", "p")

```

```{r InvarianceTab, results='asis'}
apa_table(Invariance.table, align = "c",  caption = "Measurment Invariance analysis on CFA sample (n=262) across native and non-native English speakers.", note = " a = Metric vs Configural; b = Scalar vs Metric; c = Residual vs Scalar; * =  df of model comparison", landscape = T, font_size = "footnotesize" )

```

```{r samplesizeIRT, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE, results='hide', width=60}
set.seed(222)
m <-  23 # item number
n <- c(50, 75, 100, 150, 200, 300, 350) # sample size
design <- as.data.frame(n)

irtGenerate <- function(condition, fixed_objects = FALSE) {
n <- condition$n

a <- matrix(rlnorm(m,.2,.3))
diffs <- t(apply(matrix(runif(m*4, .3, 1), m), 1, cumsum))
diffs <- -(diffs - rowMeans(diffs))
diffs<- diffs + rnorm(m)
d <- diffs
dat <- simdata(a, d, n, itemtype = 'graded')
return(dat)
}

irtAnalyze <- function(condition, dat, fixed_objects = NULL) {
mod <- mirt(dat, 1, itemtype = 'graded', verbose = FALSE)
simpars <- coef(mod, simplify = TRUE, digits = Inf)$items
irtpars <- c(a = simpars[,1], d = simpars[,2])
return(irtpars)
}

irtSummarize <- function(condition, results,
fixed_objects = NULL) {
apop <- fixed_objects['alpha', ]
dpop <- fixed_objects['d', ]
simrmse <- RMSE(results, c(apop, dpop))
out <- c(RMSE = simrmse)
return(out)
}

simres <- runSimulation(design, replications = 100,
parallel = TRUE, generate = irtGenerate,
analyse = irtAnalyze, summarise = irtSummarize,
packages = c('mirt'))
simres

colind <- grep(".a.", colnames(simres))
sima <- as.data.frame(simres[, colind])
nvec <- as.numeric(simres$n)

meanRMSE <- rowMeans(sima)
names(meanRMSE) <- n
round(meanRMSE, 2)
matplot(nvec, log(sima), type = "l", col = 1, lty = 1,
ylab = "log(RMSE)", xlab = "sample size",
main = "Graded Response Model", xaxt = "n")
axis(1, at = nvec)
```

To gather more information on our retained one-factor solution, we sought Item Response Theory (IRT). IRT complements the conventional classical test theory-based analysis by gathering information on item discrimination and item difficulty. IRT judges an item's quality by providing item information in the light of participants' trait level ($\theta$). We gathered evidence on item quality as well as item fit, person fit and model by fitting a graded response model  in RStudio with the "mirt" package [@R-mirt]  [@R-mirt]. We did a Monte Carlo simulation using "SimDesign" package [@R-SimDesign]  with sample sizes varying from 50-350 and calculated average root mean squared error(RMSE) to estimate the optimal sample size for the graded response model with 23 items. The RMSE became stable for n = 200 to 300 (RMSE ranging between .25-.35). Our sample size within the estimated sample size for stability.

```{r, include =F}
irt.data <- na.omit(irt.data)
irt.data.1 <- irt.data %>% 
  dplyr::select( CS03, CS04, CS06, CS07,CS08,CS10,CS11, CS12, 
                 CS13,CS14,CS15,CS17, CS18, CS19, CS20 ,CS21, CS22)

fit <- mirt(irt.data.1, model = 1, itemtype = 'graded', 
               SE = TRUE, Se.type = 'MHRM',
               technical = list(NCYCLES = 10000))

# Model Parameters ####
params <- coef(fit, IRTpars = TRUE, simplify = TRUE, rawug = FALSE) 
items <- data.frame(params$items)
se <- coef(fit, printSE = TRUE)

# Model Fit (degrees of freedom too low to check model fit)
model <- M2(fit)

#Item fit
item.fit<- itemfit(fit, fit_stats = c("S_X2", "G2","Zh", "infit"),
                                  impute=10)

#Identifying Missfit items based on RMSEA Value
item_misfits <- subset(item.fit, RMSEA.S_X2 >= .06)

## Creating IRT table
item.fit2 <- item.fit[,c(4,6,11,12,13,14)] 
colnames(item.fit2) <- c("Standardized Outfit", "Standardized Infit", "S-Chi-square", "df", "RMSEA", "p")

IRT.details <- cbind(items ,item.fit2)

IRT.details <- IRT.details %>% 
 dplyr::mutate_if(is.numeric, round, digits=3)

IRT.details <- tibble::rownames_to_column(IRT.details, "Items")
#write_csv(IRT.details, "table_raw/IRT_parameters.csv")

#-3<b<3
#.5<a<2; baker p12

mean.discrimination <- mean(IRT.details$a)
sd.discrimination <- sd(IRT.details$a)
range.descrimination <- range(IRT.details$a)

cat.prob <- IRT.details[,c(3:6)]
cat.prob.long <- reshape::melt(cat.prob)
range.difficulty <- range(cat.prob.long$value)

range.outfit <-  range(IRT.details$Outfit)
range.infit <-  range(IRT.details$Infit)
range.rmsea <- range(IRT.details$RMSEA)

# Person Fit ####

personfit <- personfit(fit) 
personfit_model_misfits <- subset(personfit, Zh < -2)
rownames(personfit_model_misfits)
nrow(personfit_model_misfits)
hist(personfit$Zh, xlab = "Zh Statistics",main = "F1 Person Fit")
abline(v = -2, lwd = 2, lty = 2)

person <- ggplot(personfit, aes(x=Zh)) + geom_histogram(binwidth=.5,col=I("black"),fill="palegreen4")+geom_vline(xintercept = -2)+
  labs( y="Number of Participants", x = "Zh Statistics")+apatheme + xlim(-4,2)+
   scale_y_continuous(limits = c(0, 40), expand = expand_scale(mult = c(0, 0.1)))

# annotation.1 <- data.frame(
#    x = (-3.1),
#    y = (20),
#    label = ("Most of the participants
#         are within the guidline")
# )
# 
# per.anno <-person + geom_label(data=annotation.1, aes( x=x, y=y, label=label),                 , 
#            color="orange", 
#            size=5 , angle=45, fontface="bold" )

ggsave("Figures/person.png",person , width = 12, height = 8, dpi = 600, bg = "white")

# Plots ####

OCC <- plot(fit, type = "trace", facet =T, main = "Communication Scale OCC")
info <- plot(fit, type = "infoSE", main = "Communication Scale")
iteminfo <- plot(fit, type = "infotrace", 
                    facet_items = T, main = "Communication Scale")

#conditional reliability
reliability <- plot(fit, type = 'rxx', theta_lim = c(-6, 6), 
      main="" )

marginal.rel <- marginal_rxx(fit)

#scale characteristics curve
scale <- plot(fit, type = 'score', theta_lim = c(-6, 6), main = "")

#-3<b<3
#.5<a<2; baker p12
```

```{r OCC, include=F}
ggicc <- function(model, item, theta){
  Theta <- matrix(seq(-theta,theta, by = .1))
  iteminfo <- mirt::extract.item(model, item)
  P <-mirt::probtrace(x=iteminfo , Theta=Theta  )
  icc <- data.frame(P = P, Theta = Theta)
  colnames(icc) <- c(paste("P", 1:ncol(P), sep=''), "Theta")
  icc2<- reshape(icc, direction='long', varying = paste("P", 1:ncol(P), sep=''), v.names = 'P',
                 times = paste("P", 1:ncol(P), sep=''))
  plot <- ggplot2::ggplot(icc2,aes(Theta,P, col =time))+geom_line()+xlab(expression(theta)) + 
    ylab(expression(P(theta)))+ theme(legend.title=element_blank())
  return(plot)

}

library(ggsci)

item03 <- ggicc(fit,1,6)+apatheme+ggtitle("item03")+apatheme+scale_color_npg()+
  scale_y_continuous(breaks=c(0,.5,1), limits = c(0, 1))

item04 <- ggicc(fit,2,6)+apatheme+ggtitle("item04")+apatheme+scale_color_npg()+
  scale_y_continuous(breaks=c(0,.5,1), limits = c(0, 1))

item06 <- ggicc(fit,3,6)+apatheme+ggtitle("item06")+apatheme+scale_color_npg()+
  scale_y_continuous(breaks=c(0,.5,1), limits = c(0, 1))

item07 <- ggicc(fit,4,6)+apatheme+ggtitle("item07")+apatheme+scale_color_npg()+
  scale_y_continuous(breaks=c(0,.5,1), limits = c(0, 1))

item08 <- ggicc(fit,5,6)+apatheme+ggtitle("item08")+apatheme+scale_color_npg()+
  scale_y_continuous(breaks=c(0,.5,1), limits = c(0, 1))

item10 <- ggicc(fit,6,6)+apatheme+ggtitle("item10")+apatheme+scale_color_npg()+
  scale_y_continuous(breaks=c(0,.5,1), limits = c(0, 1))

item11 <- ggicc(fit,7,6)+apatheme+ggtitle("item11")+apatheme+scale_color_npg()+
  scale_y_continuous(breaks=c(0,.5,1), limits = c(0, 1))
ggsave("Figures/ideal_icc.png",item11 , width = 12, height = 8, dpi = 600, bg = "white")

item12 <- ggicc(fit,8,6)+apatheme+ggtitle("item12")+apatheme+scale_color_npg()+
  scale_y_continuous(breaks=c(0,.5,1), limits = c(0, 1))

item13 <- ggicc(fit,9,6)+apatheme+ggtitle("item013")+apatheme+scale_color_npg()+
  scale_y_continuous(breaks=c(0,.5,1), limits = c(0, 1))

item14 <- ggicc(fit,10,6)+apatheme+ggtitle("item14")+apatheme+scale_color_npg()+
  scale_y_continuous(breaks=c(0,.5,1), limits = c(0, 1))

item15 <- ggicc(fit,11,6)+apatheme+ggtitle("item15")+apatheme+scale_color_npg()+
  scale_y_continuous(breaks=c(0,.5,1), limits = c(0, 1))

item17 <- ggicc(fit,12,6)+apatheme+ggtitle("item17")+apatheme+scale_color_npg()+
  scale_y_continuous(breaks=c(0,.5,1), limits = c(0, 1))

item18 <- ggicc(fit,13,6)+apatheme+ggtitle("item18")+apatheme+scale_color_npg()+
  scale_y_continuous(breaks=c(0,.5,1), limits = c(0, 1))

item19 <- ggicc(fit,14,6)+apatheme+ggtitle("item19")+apatheme+scale_color_npg()+
  scale_y_continuous(breaks=c(0,.5,1), limits = c(0, 1))

item20 <- ggicc(fit,15,6)+apatheme+ggtitle("item20")+apatheme+scale_color_npg()+
  scale_y_continuous(breaks=c(0,.5,1), limits = c(0, 1))

item21 <- ggicc(fit,16,6)+apatheme+ggtitle("item21")+apatheme+scale_color_npg()+
  scale_y_continuous(breaks=c(0,.5,1), limits = c(0, 1))

item22 <- ggicc(fit,17,6)+apatheme+ggtitle("item22")+apatheme+scale_color_npg()+
  scale_y_continuous(breaks=c(0,.5,1), limits = c(0, 1))

ICC <- cowplot::plot_grid(item03, item04, item06, item07, item08, 
                   item10, item11, item12, item13, item14, item15,  item17, item18, item19,
                   item20, item21, item22,
                      labels = "AUTO",
                     ncol=3,  
                    label_size = 15, 
                     align = "v")

ggsave("Figures/ICC.png",ICC , width = 12, height =14, dpi = 600, bg = "white")

```

```{r figICC, eval=FALSE, fig.cap="Option Charecteritics Curve.", warning=FALSE, include=FALSE, out.height='100%'}

knitr::include_graphics("Figures/ICC.png", dpi =600)

```

```{r item-plots, inclide =F}

iteminfo <- plot(fit, type = "infotrace",which.item =2)

itemplot.01 <- tabledown::ggiteminfo(fit,1,6)+apatheme+
  ggtitle(" Item03")+theme(legend.position = "none")+
  labs(y = "Item Information")+
  scale_color_npg()+geom_hline(yintercept = .20, colour ="red")+geom_area(fill ="palegreen4" )+
  xlab(expression(theta))+ ylab(expression(I(theta)))

itemplot.02 <- tabledown::ggiteminfo(fit,2,6)+apatheme+
  ggtitle(" Item04")+theme(legend.position = "none")+
  labs(y = "Item Information")+
  scale_color_npg()+geom_hline(yintercept = .20, colour ="red")+geom_area(fill ="palegreen4" )+
  xlab(expression(theta))+ ylab(expression(I(theta)))

# good.info <- cowplot::plot_grid( item6.iic1, item7.iic1, item11.iic1, item12.iic1, item14.iic1, item15.iic1 , item18.iic1, item19.iic1, item20.iic1, item21.iic1, item22.iic1, 
#                                  labels = "AUTO",ncol=3, label_size = 15,align = "v")
# 
# 
# ggsave("Figures/good.info.png",good.info, width = 12, height = 8, dpi = 600, bg = "white")

test.info <- tabledown::ggtestinfo(irt.data.1, fit)+ apatheme+geom_line(colour = "black", size=1.5)+geom_area(fill ="#E64B3599" )
# +geom_vline(xintercept = -3, colour ="black")+
#   geom_vline(xintercept = .5, colour ="black")
ggsave("Figures/TIC.png",test.info , width = 6, height = 4, dpi = 600, bg = "white")

library(plotly)
info.p <- plotly::ggplotly(test.info)

```

```{r figTIC, fig.cap = "TIC.",  out.height='100%', warning=FALSE}

knitr::include_graphics("Figures/TIC.png", dpi =600)

```

```{r conrel-print, include=F}
##
png(filename="Figures/conrel.png", 
    type="cairo",
    units="in", 
    width=12, 
    height=2, 
    pointsize=12, 
    res=300)
library(latticeExtra)
library(directlabels)
font.settings <- list(
  font = 1,
  cex = 6,
  fontfamily = "sans")

apa.latice  <- list(
  par.xlab.text = font.settings,
  par.ylab.text = font.settings,
  axis.text = font.settings,
  sub.text = font.settings,
  add.text = font.settings)

 plot(fit, type = 'rxx', theta_lim = c(-6, 6), 
      main="Conditional Reliability Plot", par.settings=c(apa.latice),col.line = "#E69F00",lwd = 3.5 )

dev.off()
```

```{r figconrel, fig.cap = "Conditional Reliability.",  out.height='100%', warning=FALSE}

knitr::include_graphics("Figures/conrel.png", dpi =600)

```

Marginal reliability is based on the true score model (Lord & Novick, 1968) and is an estimate of the overall reliability of a test based on the average conditional standard errors. Often it is close in value to coefficient alpha (and sometimes it may even be identical). Alpha provides a lower estimate of marginal reliability.

```{r dif, eval=FALSE, include=FALSE}
library("lordif")
gender <- as.data.frame(data$Gender)
colnames(gender) <- "Gender"
gender$Gender <- as.factor(gender$Gender)
irt.data <- na.omit(irt.data)
communication <- sem.data[,c(2:24)]
dif.data <- cbind(gender, communication)
dif.data <- na.omit(dif.data)
dif.data.full <- as.data.frame(dif.data)
DIF <- lordif(as.data.frame(irt.data), group = dif.data.full$Gender, criterion = "Chisqr")

DIF$stats[,1:5]
plot(DIF)
```

```{r}

val.data<- validity.data.full %>% 
  dplyr::select( CS03, CS04, CS06, CS07,CS08,CS10,CS11,    CS12, 
                 CS13,CS14,CS15,CS17, CS18, CS19, CS20 ,CS21, CS22) %>% 
  rowwise() %>% 
  mutate(Total_communication = sum(across(starts_with("CS")), na.rm = T),
         F1 = sum(CS15,CS22,CS20,CS14,CS18,CS13,CS06,CS10),
         F2 = sum(CS19,CS04,CS17,CS12,CS08,CS07))

val.data.2 <- as.data.frame(cbind(val.data$Total_communication, validity.data.full$Total_HS, validity.data$Total_LS, validity.data.full$Total_SE))

colnames(val.data.2) <- c("Communication", "Hopelessness", "Life Satisfaction", "SE")
validity <- na.omit(val.data.2)
```

```{r correlationfunc, include =F}
glrstab<- function(x, export=FALSE) {

 r <-corr.test(x)$r	#taking just the correlation matrix; no N, or p
 p <-corr.test(x)$p	#taking the p*s

#define notions for significance levels
 mystars <- ifelse(p < .001, "**"
                   , ifelse(p < .01, "**"
                            , ifelse(p < .05, "*"
                                     , ifelse(p < .10, "+", " "))))

 #round r, define new matrix Rnew with the correlations from rnd and paste mystars
 rnd  <- papaja::printnum(r, gt1 = FALSE, digits = 2)  #round, drop leading 0 - Thanks CRSH!								                     
 Rnew <- matrix(paste(rnd, mystars, sep=""), ncol=ncol(rnd))

  #remove 1.0 correlations from diagonal  and set the strings
 diag(Rnew) <- ''		
 Rnew[upper.tri(Rnew)] <- ''

 rownames(Rnew) <- paste(1:ncol(rnd), colnames(rnd), sep=" ")         #define number and name
 colnames(Rnew) <- paste(1:ncol(rnd), "", sep="") 			       #define number

#fun-part: we trim the top half 
 Rnew[upper.tri(Rnew)] <- ''			
 Rnew

 #Rnew <- cbind(round(describe(x)[,3:4],2), Rnew)		     #describe x, M sD - put them in the matrix
 #colnames(Rnew)[1:2] <- c("M","SD")					      		#Beschriftung der neuen Spalten
 Rnew <- Rnew[,1:(ncol(Rnew)-1)]							        	#delete the last column (ugly)

 #export to clipboard

   if (export==TRUE){
   result<-write.table(Rnew
                       , "clipboard"
                       , sep=";"
                       , row.names=FALSE)
 }
 else result <- Rnew
 return(result)

}

```

```{r correlations, include =F}
Cor <-glrstab(validity) #the function in action!

cor.tab <- as_tibble(cor(validity))

```

```{r tabvalidity,  results='asis'}
papaja::apa_table(Cor, escape  = FALSE
                , caption = "Correlation matrix of the main variables"
                , note    = "**p < .001")
```

# Results

# Discussion

## Ethical Consideration

All procedures performed in studies involving human participants were in accordance with the 1964 Helsinki declaration and its later amendments or comparable ethical standards. This article does not contain any studies with animals performed by any of the authors.

## Data and code availability
All code and data underlying this article is available on a public GitHub repository (https://github.com/masiraji/Communication).

\newpage

# References

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id="refs" custom-style="Bibliography"></div>
\endgroup