### 2022-02-04 manuscript.Rmd
EFA results updated

---
title             : "Psychometric Evaluation of the Bangla Communication Scale Among Adolescents:Latent structure, Reliability, validity, and gender invariance in a Bangladeshi adolescent sample"
shorttitle        : "Bangla Communication Scale"

author: 
  - name          : "Nusrat Jahan"
    affiliation   : "1"
    corresponding : no    # Define only one corresponding author
    role:         
      - Conceptualization
      - Project Management
      - Data Curation
      - Writing - Original Draft Preparation
  - name          : "Mushfiqul Anwar Siraji"
    affiliation   : "2"
    corresponding : no   # Define only one corresponding author
    address       : "Monash University Malaysia,Jalan Lagoon Selatan, 47500 Bandar Sunway,Selangor Darul Ehsan, Malaysia."
    email         : "mushfiqul.anwarsiraji@moansh.edu"
    role:
      - Conceptualization
      - Project Management
      - Formal Analysis & Data Visualization
      - Writing - Original Draft Preparation; Review & Editing

  - name          : "Zinnatul Borak"
    affiliation   : "1"
    corresponding : yes   # Define only one corresponding author
    address       : "Department of Educational and Counselling Psychology,University of Dhaka,Dhaka 1000"
    email         : "institutional email of Bonhee mam"
    role:
      - Conceptualization
      - Writing - Review & Editing

affiliation:
  - id            : "1"
    institution   : "Department of Educational and Counselling Psychology,University of Dhaka"
  - id            : "2"
    institution   : "Monash University, Department of Psychology, Jeffrey Cheah School of Medicine and Health Sciences, Malaysia"

authornote: |
  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

  Enter author note here.

abstract: |
  One or two sentences providing a **basic introduction** to the field,  comprehensible to a scientist in any discipline.

  Two to three sentences of **more detailed background**, comprehensible  to scientists in related disciplines.

  One sentence clearly stating the **general problem** being addressed by  this particular study.

  One sentence summarizing the main result (with the words "**here we show**" or their equivalent).

  Two or three sentences explaining what the **main result** reveals in direct comparison to what was thought to be the case previously, or how the  main result adds to previous knowledge.

  One or two sentences to put the results into a more **general context**.

  Two or three sentences to provide a **broader perspective**, readily comprehensible to a scientist in any discipline.

  <!-- https://tinyurl.com/ybremelq -->

keywords          : "keywords"
wordcount         : "X"

bibliography      : ["r-references.bib", "references.bib"]

floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no

documentclass     : "apa6"
classoption       : "man"
output            : 
  papaja::apa6_pdf:
    keep_tex: false
    latex_engine: xelatex
    includes:
      in_header: header.tex
mainfont: Arial
---

```{r setup, warning=FALSE, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = F, message = F, dpi=600, fig.width = 6,
 fig.asp = 0.8,out.width = "80%", dev=c('png','postscript'))
#options(knitr.duplicate.label = "allow")
set.seed(123)
par(family = "Arial")
```

```{r pacman, eval=FALSE, include=FALSE}
#You need to run this chunk 1X in your system 
#install.packages("pacman")#Run this if you don't have pacman installed.
pacman::p_load(MOTE, tidyverse, psych, lavaan,kableExtra,gt,gtsummary, mirt,likert,kutils,semPlot,semTable,semTools,ggcorrplot,dlookr,qgraph,paran,EFA.MRFA,VIM,DiagrammeR,DiagrammeRsvg,ggplot2,cowplot,rsvg,questionr,magick, simsem,readxl, stringr,RColorBrewer,WrightMap, foreign,DT,SimDesign)
library(papaja)
pacman::p_install_gh("jthomasmock/gtExtras")
pacman::p_install_gh("crsh/papaja")
pacman::p_install_gh("masiraji/tabledown")
pacman::p_install_gh("crsh/citr")

papaja::r_refs("references.bib")
#if you dont have LaTeX installed run the following lines
#install.packages("tinytex")
#tinytex::install_tinytex()
renv::snapshot()
```

```{r library,include=FALSE}
#This chunk holds code for calling the required packages

library(papaja) # devtools::install_github("crsh/papaja")
library(lavaan)
library(semPlot) #devtools::install_github('SachaEpskamp/semPlot',  dependencies = T)
library(semTools)
library(MOTE)
library(car)#reversecoding 
library(psych)
library(dlookr)
library(plyr)
library(tidyverse)
library(qgraph)
library(kableExtra)
library(paran) # Parallel analysis
library(EFA.MRFA) # Hull method
library(DiagrammeR) #devtools::install_github('rich-iannone/DiagrammeR')
library(DiagrammeRsvg) #For DiagrammeR
library(rsvg) #For DiagrammeR
library(ggcorrplot)
library(semTable)
library(magick)
library(mirt)
library(gtsummary)
library(gt)
library(gtExtras)
library(likert)
library(VIM) #Missing data
library(kutils)
library(simsem)
library(tabledown)
library(readxl)
library(koRpus)# stable release
library(koRpus.lang.en)
library(reshape)
library(ggsci)
renv::snapshot()
r_refs("references.bib")
```

---
nocite: |
  `r papaja::cite_r("references.bib")`

---

Communication is a complex behaviour of exchanging information among individuals [@tanner2006advanced]. Communication plays a central role among adolescents in developing self-identity, social relationships and creates the foundation of collective social activity [@haslett1989children; @conti2008emotional; @spencer2013language]. Inadequate communication skill may cause poor peer relationship resulting long-term socio-emotional difficulties including social anxiety, stress, low self-esteem and poor academic performance [@brinton2004social; @reed2020relative].

Often adults picture adolescents having inadequate and inept communication skills [@thurlow2003teenagers;@stern2005self]. Media representation of adolescents often includes "storm-and stress", self absorbed and disengaged type behaviours [@porteous1980adolescents;@stern2005self]. As such adolescents are often labelled as "lazy" and "disrespectful" by the adults [@public1999kids]. On the contrary adolescents are highly engaged in work, community services and extracurricular activities and also more aspiring to earn an college degrees [@debard2004millennials; @schneider1999ambitious]. Such a discrepancy between the reality of adolescent's image and adult's perception of the adolescents might be attributed to the mismatch of communication skills. The communication pattern of adolescents might not necessarily same as the adults. In addition to face to face communication, adolescents are vastly exposed to different virtual communication platforms. This may cause them to face more complex social challenges than the adults [@thurlow2003teenagers]. "Communication capital" expresses the potential of civic-engagement that incorporates developing social relationships and influences collective social activity. The more communication capital an individual has the easier the instances of civic engagement become.

Understanding the adolescents' communication skill  vital as it is considered as the "key skill" in the education [@thurlow2001talkin] and employment market [@olszewski2017training]. In the western society, adolescents are now facing high unemployment [@lindsay2014employers]. Lack of adequate communication skill is one of the root causes of this high unemployment [@lindsay2014employers]. Similarly, lack of proper communication skill often promotes the propensity of anti-social behaviours and risk of exclusion from schools [@clegg2009language; @conti2004social].

To promote better understanding of subject contents assessing the communication skill among adolescents is highly required. For this purpose "Communication Skill" sub-skill set [@barkman2002four]  was developed in 2002 as a part of The  National  On-line  Youth  Life  Skills  Evaluation System [@mincemoyer2005measuring] and since then it has been extensively used [@fitzpatrick2005life].

# Study-1: Translations and Exploratory Factor Analysis
Study-1 had two objectives. First we translated the 23 items of the Communication Scale [@barkman2002four] from English to Bangla. Second, we conducted an Exploratory Factor Analysis (EFA) to identify the latent structure of the scale.

## Methods

### Participants

```{r data, include=F}
# data <- readxl::read_excel("Processeddata/communication_data_n_582.xlsx")
# readr::write_rds(data, "Processeddata/communication_data_n_582.rds")
data <- readr::read_rds("Processeddata/communication_data_n_582.rds")

efa.data.full <- subset(data, Sample == "EFA")
cfa.data.full <- subset(data, Sample == "CFA"| Sample=="Validity")
validity.data.full <- subset(data, Sample == "Validity")

sem.data <- data[,c(2,6:66)]
efa.data <- subset(sem.data , Sample == "EFA")
efa.data <- efa.data[,c(2:24)]
efa.data <- na.omit(efa.data)

cfa.data<- subset(sem.data , Sample == "CFA"| Sample=="Validity")
cfa.data <- cfa.data[,c(2:24)]
cfa.data <- na.omit(cfa.data)

validity.data <- subset(sem.data , Sample == "Validity")
validity.data <-na.omit(validity.data)

irt.data <- sem.data[,2:24]
```

```{r sescrptives, include=FALSE}
desc <- efa.data.full[,c(3,4,5)]
## Recoding desc$Sex
desc$Gender <- fct_recode(desc$Gender,
  "Boy" = "male",
  "Girl" = "female"
)
colnames(desc) <- c("Gender", "Age", "Socio Economic Status")

des_tab <- desc %>%
  tbl_summary(
    by = Gender,
    statistic = list(all_continuous() ~ "{mean} ({sd})",
                     all_categorical() ~ "{n} ({p}%)"),
    type = list(Age ~ 'continuous'),
    digits = all_continuous() ~ 2,

    missing = "no"
    ) %>% add_overall() %>% bold_labels() %>% 
modify_header(label ~ "**Variable**")  %>%
  modify_caption("Demographic Characteristics")

as_tibble(des_tab) -> desc_tibble 
as_kable_extra(des_tab, format = "latex",booktabs = T) -> desc_kable #save it as a knitr::kable

des.2 <- desc %>% 
  psych::describeBy("Gender")
female <- as.tibble(des.2$Girl)
male <- as.tibble(des.2$Boy)

```

A cross-sectional survey was used to collect data from a large sample of students of grade 8-12 (n = `r nrow(efa.data)`) from 8 schools following convenience sampling method.One participant was excluded due to incomplete data.  Participants were recruited following convenience sampling technique. For exploring the initial factor structure the recommended sample size is 250-300 [@comreyFirstCourseFactor1992; @schonbrodtWhatSampleSize2013]. Among `r nrow(efa.data)` participants, `r female[1,2]` were female aged between `r female[2,8]` to `r female[2,9]` years (`r round(female[2,3],2)`±`r round(female[2,4],2)`). `r male[1,2]`  were male with an age range between `r male[2,8]` participants `r male[2,9]` years (`r round(male[2,3],2)`±`r round(male[2,4],2)`). `r desc_tibble [4,2]` participants belonged to middle socio-economic status. `r desc_tibble [3,2]` and `r desc_tibble [5,2]` participants belonged to lower and upper socio-economic status-respectively.

### Materials

#### Communication Scale

Communication Scale [@barkman2002four] is a sub-skill set of The  National  On-line  Youth  Life  Skills  Evaluation System [@mincemoyer2005measuring]. It has 23 items with a 5 Point Likert Type response scale (0=Never, 1=Rarely,  2=Sometimes,  3=Often,  4=Always). The total score range is 0-92 where a higher score would indicate higher communication skills among adolescents (age range 12-18). The internal consistency of the total scale, Cronbach $\alpha$ was .79

#### Bangla Communication Scale
We followed the International Test Commission  guidelines [@bartramITCGuidelinesTranslating2018] while translating the Communication Scale in Bangla. At first two bilingual researchers did the forward translation. These two forward translations were synthesized by the authors. Another four bilingual researchers did the backward translation of the synthesized Bangla Communication Scale. The authors again synthesized the back-translations, compared it with original scale and made necessary amendments.

### Data collection
The project received institutional ethics clearance from Department of Educational and Counselling Psychology (Project ID: ---). Prior to data collection necessary authorization from school's authority and assent from the participating were obtained. Necessary explanations were given in oral and written forms.It  was also mentioned in the explanatory statement that their participation was voluntary and that they could withdraw from participation any time without being penalized.  Data collection was commenced between November 2021 to January 2022. The data collection took place in the classroom where students were at first briefed about 'communication skill'. Next. they filled up their socio-demographics information and responded to our Bangla Communication Scale. All personal information (name, school,class) was codified and encrypted, producing a anonymous database.

### Analytic strategies
We used R (version 4.1.0) [@R-base], including several R packages [@R-lavaan; @R-mirt; @R-psych; @R-tabledown], for our analyses. Prior to the data analysis normality assumptions were checked. Our data violeted both univariate and multivariate normality. As such we used polychoric correlation matrix and 'principal axis" factor extraction method to conduct the exploratory factor analysis [@watkinsStepbyStepGuideExploratory2020]. An oblique rotation method was employed. A posterior sampling adequacy was estimated using KMO statistics [@kaiserIndexFactorialSimplicity1974]. TO identify the optimum number of factors required to explain the latent structure of our scale we used Scree plot [@cattellScreeTestNumber1966a], Horn's parallel analysis [@hornRationaleTestNumber1965], Hull method [@lorenzo-sevaHullMethodSelecting2011] and Minimum average partials method (MAP) [@velicerDeterminingNumberComponents1976]. Lastly, to identify the simple structure, we followed the following guidelines  (i) no factors with fewer than three items (ii) no factors with a factor loading <0.3 (iii) no items with cross-loading greater than .3 across factors [@childEssentialsFactorAnalysis2006a; @mulaikFoundationsFactorAnalysis2009; @watkinsStepbyStepGuideExploratory2020]

### Results & Discussion
```{r, gtfig, include =F}

gt.data <- na.omit(efa.data)

names(gt.data)[names(gt.data) == "RCS02"] <- "CS02"
names(gt.data)[names(gt.data) == "RCS05"] <- "CS05"

#Creating long data
gt.data.long <- as.data.frame(gather(gt.data, Items, value))
gt.data.long$value <- as.numeric(as.character(gt.data.long$value))

gt.data.tab <- gt.data.long %>% 
  group_by(Items) %>% 
  # calculate summary stats & create data for the histogram and density plot
  dplyr::summarise(
    nr = n(),
    mean = mean(value, na.rm = TRUE),
    med = median(value, na.rm = TRUE),
    sd = sd(value, na.rm = TRUE),
    hist_data = list(value),
    dens_data = list(value),
    .groups = "drop"
  ) %>% 
  gt()

gt.likert <-  as.data.frame(gt.data)

#gt.likert <-  mutate(gt.likert, across(starts_with("ASQ"), ~unname(recod[.])))

gt.likert.Factor = as.data.frame(lapply(gt.likert,factor,
                                     ordered = T))

#get the items name
gt.items <- names(gt.likert.Factor) 
#Calculate percentage
gt.percentage <- kutils::likert(gt.likert.Factor, vlist = gt.items )

gt.percentage <- gt.percentage$table %>% 
  as.data.frame(.)

#data wrangling
labels <- c("never", "rarely","sometimes", "often", "always","Total")
as.data.frame(labels)

full.percentage <- cbind(labels,gt.percentage) #tables with labels  
full.percentage<- t(full.percentage ) #transpose
as.data.frame(full.percentage)
full.percentage1 <- full.percentage[-1,-6] #removing 1st row and total column
#full.percentage2 <- full.percentage1[, c(3, 4, 1, 2,5)]# rearranging
as.data.frame(full.percentage1)

colnames(full.percentage1) <- c("Never", "Rarely","Sometimes", "Often", "Always")

Items <- rownames(full.percentage1)
as.data.frame(Items)
full.percentage3 <- cbind(Items,full.percentage1)

full.percentage3 <- full.percentage3[order(Items),] 
full.percentage3 <- as.data.frame(full.percentage3[,-1])

full.percentage3 <- full.percentage3%>% 
  gt()

gt.table <- gt.data.tab$'_data'
gt.liket <- full.percentage3$"_data"

Table <-  cbind( gt.table, gt.liket )%>% 
  gt()%>% 
  # histogram and density plots
  gtExtras::gt_sparkline(
    hist_data,
    type = "histogram",
    line_color = "black", 
    fill_color = "#00A08799",
    bw = 1,
    same_limit = TRUE)%>%
  gtExtras::gt_sparkline(
    dens_data,
    type = "density",
    line_color = "black", 
    fill_color = "#E64B3599",
    bw = 0.75,
    same_limit = TRUE
  )%>%
  # format decimals
  fmt_number(columns = mean:sd, decimals = 1) %>%
  # header
  tab_header(
    title = md("**Communication**"),
    subtitle = md("Summary Descriptives")
  ) %>% 
  tab_footnote(
    footnote = md("**Histogram**"),
    locations = cells_column_labels(columns = hist_data)
  ) %>%
  tab_footnote(
    footnote = md("**Density**"),
    locations = cells_column_labels(columns = dens_data)
  ) %>% 
  #create groups of columns
  tab_spanner(
    label = "Items",
    columns = Items
  ) %>%
  tab_spanner(
    label = "Summary Statistics",
    columns = nr:sd
  ) %>%
  tab_spanner(
    label = "Graphics",
    columns = hist_data:dens_data
  )  %>%
  # change column names to appear in the table
  cols_label(
    Items = ("Items"),
    nr = ("n"),
    mean = ("Mean"),
    med = ("Median "),
    sd = (("SD")),
    hist_data = "Histogram",
    dens_data = "Density"
  ) %>%

  tab_spanner(
    label = "Response Pattern",
    columns = "Never":"Always"
  ) %>%
  # set alignment as per wish
  cols_align(
    align = "center",
    columns = everything()
  ) %>%
  opt_align_table_header(align = "left") %>%
  gtExtras::gt_plt_dot(
    mean,
    Items,
    palette = "ggthemes::fivethirtyeight")

gtsave(Table, "Figures/efa_communication.png",vwidth = 7000)

```

```{r ggplot2, include=F}
#This chunk holds code for creating ggplot2 based apatheme for plots.
apatheme=theme_bw()+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        #text=element_text(family = "Helvetica"),
        axis.text.x = element_text(size = 15),
        axis.text.y = element_text(size = 15),
        axis.title.x = element_text(size = 15),
        axis.title.y = element_text(size = 15),
        plot.title = element_text(size = 15),
        legend.text = element_text(size = 15),
        legend.title=element_blank(),
        axis.line.x = element_line(color='black'),
        axis.line.y = element_line(color='black'),
        panel.border = element_rect(color = "black",
                                    fill = NA,
                                    size = 1))

```

```{r EFAdata, include=FALSE}

#library(VIM)
missing <- VIM::aggr(efa.data, plot =T)

```

```{r item-analysis, include=FALSE, results='asis'}
Cron.bach <- psych::alpha(correlations$rho,check.keys=TRUE) 
Cron.bach$total %>% 
  knitr::kable(caption = "Alpha", digits = 2) 
Cron.bach$alpha.drop %>% 
  knitr::kable(caption = "Alpha value if item is dropped", digits = 2) 
Cron.bach$item.stats %>% 
  knitr::kable(caption = "Corrected Item total Correlation", digits = 2)

low.r.corec <- (min(Cron.bach$item.stats$r.cor))  #minimum item total correlation
high.r.corec <- (max(Cron.bach$item.stats$r.cor))     # maximum item total correlation

ordinal.alpha <- Cron.bach$total$raw_alpha
ordinal.alpha.drop <- as.data.frame(Cron.bach$alpha.drop$raw_alpha)
minimum.ordinal.alpha.drop <- min(ordinal.alpha.drop )
maximum.ordinal.alpha.drop <- max(ordinal.alpha.drop )
```

```{r EFAassumptions, include=FALSE}

#KMO test
KMO <- psych::KMO(efa.data)

# Test of correlation matrix
bartlet <- psych::cortest.bartlett(efa.data, n =300)

decriptives <- tabledown::des.tab(efa.data)
decriptives <- cbind(decriptives, ordinal.alpha.drop )
colnames(decriptives) = c("Items","Mean", "SD", "Skew", "Kurtosis", "Shapiro-Wilk Statistics","Item-Total Correlation", "Alpha Drop")
descriptives <- cbind()

item.total.correlation <- range(decriptives$`Item-Total Correlation`)
```

#### Sampling adequacy
Kaiser-Meyer-Olkin (KMO) [@kaiserIndexFactorialSimplicity1974] statistics was used to check the sampling adequacy.  The overall KMO value for 23 items was `r apa(KMO$MSA,2,T)`, which was above the cut-off value of .50, indicating an adequate sample.

#### Descriptive statistics and item analysis
Table\@ref(tab:tabDes) reports univariate descriptive statistics for the 23 items. All items were skewed and violeted univariate normality assessed by  The Shapiro-Wilk test indicated all the items violated normality assumptions [@shapiro1965analysis]. Mardia's Test of multivariate normality [@mardiaMeasuresMultivariateSkewness1970] yielded Multivariate skew = 4030.49 (p <0.001) and multivariate kurtosis = 15.1 (p <0.001) indicating the violation of multivariate normality as well. As such we used polychoric correlation matrix which is more robust towards these violations. Our initial item analysis yelled internal consistency coefficient ordinal alpha = `r apa(ordinal.alpha,2,T)`. Alpha drop statistics (Table\@ref(tab:tabDes)) showed no substantial increase of ordinal alpha if any item is deleted and ranged between `r apa(minimum.ordinal.alpha.drop,T)`-`r apa(maximum.ordinal.alpha.drop,T)`. As such we subjected all items to EFA.

Supplementary Table 1 and Figure \@ref(fig:corplot) depict the inter-item correlation coefficients of BCS. Bartlett’s test of sphericity [@RN1261], $\chi^2$ (`r bartlet$df`) = `rbartlet$chisq`, p < .001 indicated the inter-item correlations are significantly different than zero. However, only `r apa(cor.per,2,T)`% of the inter-item correlation coefficients were greater than absolute value of .30 in the obtained matrix. The corrected item-total correlations range was `r min(item.total.correlation)`-  `r max(item.total.correlation)` (Table\@ref(tab:tabDes)).

```{r tab-des}
apa_table(decriptives)
```

```{r CorrMatrix, include=FALSE}
#Polychoric Correlation matrix

correlations <- psych::polychoric(efa.data)
upper <- correlations$rho[upper.tri(correlations$rho, diag = F)]

min.cor <- apa(min(abs(upper)),2,F) #minimum cor.coefficient of the matrix
max.cor <- apa(max(abs(upper)),2,F) # max. correlation coefficient of the matrix

determinets <- det(correlations$rho) 
# Calculating the percentage of correlations higher than .30

BigR = sum(correlations$rho >= abs(.30) & correlations$rho < abs(1.0), na.rm =T)/2 
totR = length(efa.data)*(length(efa.data)-1)/2
cor.per <- print (BigR/totR)*100
```

```{r corplot, include =F}
#This chunk holds code for plotting the correlation matrix

corplot <- ggcorrplot(correlations$rho, hc.order = TRUE, outline.col = "white", type = "lower", 
           ggtheme = ggplot2::theme_minimal(),tl.srt = 90,colors = c("#E64B35FF", "white", "#3C5488FF") )+
  theme( panel.grid.major = element_blank(),
         axis.text.x = element_text(size = 8),
         axis.text.y = element_text(size = 8),
         panel.border = element_rect(color = "black",
                                    fill = NA,
                                    size = 1))
#p.mat <- cor_to_p(correlations$rho, 428, method = "polychoric")
#to show pvalue in the fig add p.mat=p.mat
ggsave("Figures/corplot.png",corplot, width = 6, height = 6, dpi = 600, bg ="white")

```

(ref:figCor) Inter item polychoric correlation coefficients for the 23 items. 9.8 % inter-item correlation coefficients were higher than |.30|

```{r figCor, out.height="150%",out.width="100%", fig.align= 'center', fig.cap='(ref:figCor)', results='asis', fig.align='center'}

knitr::include_graphics('Figures/corplot.png', dpi =600)

```

```{r ParallelEFA, include=FALSE}
#This chunk holds code for factor identification method : Horn
# Parallel analysis
Horn <- paran(correlations$rho, iterations=500, centile=0, quietly=FALSE,
           status=TRUE, all=FALSE, cfa=FALSE, graph=TRUE,
           color=TRUE, col=c("black","red","blue"),
           lty=c(1,2,3), lwd=1, legend=TRUE, 
           width=640, height=640, grdevice=png, seed=0, mat=NA, n=NA)

Adjusted.EV <- data.frame(Horn$AdjEv)
Adjusted.EV$type = c('Adjusted Ev')
Adjusted.EV$num = c(row.names(Adjusted.EV))
Adjusted.EV$num = as.numeric(Adjusted.EV$num)
colnames(Adjusted.EV) = c('eigenvalue', 'type', 'num')

Random.EV <- data.frame(Horn$RndEv)
Random.EV$type = c('Random EV')
Random.EV$num = c(row.names(Random.EV))
Random.EV$num = as.numeric(Random.EV$num)
colnames(Random.EV) = c('eigenvalue', 'type', 'num')

Unadjusted.EV <- data.frame(Horn$Ev)
Unadjusted.EV$type = c('Unadjusted EV')
Unadjusted.EV$num = c(row.names(Unadjusted.EV))
Unadjusted.EV$num = as.numeric(Unadjusted.EV$num)
colnames(Unadjusted.EV) = c('eigenvalue', 'type', 'num')

parallel = ggplot(Adjusted.EV, aes(x=num, y=eigenvalue, shape=type, color=type)) +
  #Add lines connecting data points
  geom_line(colour ='#DC000099')+
    #geom_line(data = Random.EV) +
  #geom_line(data = Unadjusted.EV)+
    #Add the data points.
  geom_point(size=1.5,colour ='#3C5488FF')+
  #geom_point(data = Random.EV, size=3)+
  #geom_point(data = Unadjusted.EV, size=3)+
  scale_y_continuous(name="Eigen Value", limits=c(0, 4), breaks=seq(0,4,1)) +
   #Label the x-axis 'Factor Number', and ensure that it ranges from 1-max # of factors, increasing by one with each 'tick' mark.
  scale_x_continuous(name='Factor Number (Parallel Analysis)', limits=c(0, 23),breaks=seq(0,23,4))+
    #Add vertical line indicating parallel analysis suggested max # of factors to retain
  geom_hline(yintercept=1, linetype = 'dashed') + apatheme +theme(,legend.position = "None") 
#text = element_text(size = 25)

```

```{r ScreePlot, include=FALSE}

#This chunk holds code for factor identification method : Scree Plot

Scree <- scree(correlations$rho,factors=TRUE,pc=TRUE,main="(B)",
      hline=NULL,add=F)

FA.Scree <- data.frame(Scree$fv)
FA.Scree$type = c('Factor Analysis')
FA.Scree$num = c(row.names(FA.Scree))
FA.Scree$num = as.numeric(FA.Scree$num)
colnames(FA.Scree) = c('eigenvalue', 'type', 'num')

PA.Scree <- data.frame(Scree$pcv)
PA.Scree$type = c('Principal Component Analysis')
PA.Scree$num = c(row.names(PA.Scree))
PA.Scree$num = as.numeric(PA.Scree$num)
colnames(PA.Scree) = c('eigenvalue', 'type', 'num')

scree.plot <-ggplot(FA.Scree, aes(x=num, y=eigenvalue,color=tye, shape=type))+
   geom_line(colour ='#DC000099')+
  #geom_line(data = PA.Scree)+ 
  geom_point(size=1.5,colour ='#3C5488FF')+
  #geom_point(data = PA.Scree, size=3)+
  scale_y_continuous(name="Eigen Value", limits=c(0, 5), breaks=seq(0,5,1)) +
  scale_x_continuous(name='Factor Number (Scree Plot)', limits=c(0, 20),breaks=seq(0,20,4))+
  geom_hline(yintercept=1, linetype = 'dashed') + apatheme +theme(legend.position = "None") 
#text = element_text(size = 25

```

```{r HullEFA, include=FALSE}
#This chunk holds code for factor identification method : Hull

# HULL
EFA.MRFA::hullEFA(efa.data,extr = "ULS", index_hull = "CAF", display = TRUE, graph = T,
        details = TRUE)
hull <- ggplot2::last_plot()
Hull <- hull+ ggtitle(NULL)+xlab("Factor Number (Hull Method)")+ylab("CAF")+ aes(color = '#DC000099')+geom_point(color ='#3C5488FF',size =1)+
  apatheme +theme(legend.position = "None") +scale_fill_identity()

```

```{r MAPefa, include=FALSE}
#This chunk holds code for factor identification method : MAP
#MAP
map <- psych::VSS(efa.data, rotate = "promax", fm = 'pa', n.obs =300 )
map.map <- as.data.frame(map$map)
colnames(map.map) <- "MAP Statistic"
map.statistics <- map$vss.stats[,c(1,2,5,6,7,10,11)]
full.map <- cbind(map.map,map.statistics)

write.csv(full.map, "Table_raw/map_stat.csv")
```
####Exploratory factor analysis

To identify opttimum number of factor required to express the latent structure adequetly we at first used Horn's parallel analysis [@hornRationaleTestNumber1965]. Horn's parallel analysis offered two factor solution to the latent structure (Figure \@ref(fig:facIdFig)). However, The Scree-plot [@cattellScreeTestNumber1966a] suggested one factor solution. One factor solution was also supported by MAP method [@velicerDeterminingNumberComponents1976] and Hull method [@lorenzo-sevaHullMethodSelecting2011]. Minimum average partial (MAP) method  expects the average squared off-diagonal values of the calculated partial correlation matrix to be minimum when the correct number of factors are extracted [@velicerDeterminingNumberComponents1976]. In our data set this value reached the minimum after extracting the first factor (Supplementary Table 2). Hull methop tried to find an optimal number of factors to balance model fit and the number of parameters and offeren one factor solution in our data set (Figure \@ref(fig:facIdFig)). As a result, we tested both one-factor and two-factor solutions.

```{r factors, include=F}
#This chunk holds code for compiled factor identification plots

factor <- cowplot::plot_grid(parallel, scree.plot,Hull, 
                   labels = "AUTO",
                   ncol=1, 
                  align = "v", 
                  label_fontfamily = "sans",
                  label_fontface = "plain")
ggsave("Figures/factors.png",factor, width = 5, height = 7, dpi = 600)
```

(ref:facIdFig) Factor Identification Methods (A) Parallel analysis  indicated the optimal number of factors were two. (B) Scree plot suggested one factor. (C) Hull method indicated one factor were required to balance the model fit and number of parameters.

```{r facIdFig, fig.align= 'center', fig.cap='(ref:facIdFig)',  warning=FALSE, results= 'asis'}

knitr::include_graphics("Figures/factors.png", dpi =800)

```

```{r EFA, include=FALSE}
fa.2F.1 <- fa(r=correlations$rho, nfactors = 2, fm= "pa",rotate ="promax",
              residuals = TRUE, SMC = TRUE)
AA <- print(fa.2F.1, cut = .3, digits = 3, sort = TRUE)

reduced.model.2F.1 <- dplyr::select(efa.data ,
                                    -c( CS09, CS03, CS16, CS01, RCS05))

correlations.red.2F.1 <- polychoric(reduced.model.2F.1)

fa.2F.2 <- fa(r=correlations.red.2F.1$rho, nfactors = 2, fm= "pa",rotate ="promax",
              residuals = TRUE, SMC = TRUE)

BB <- print(fa.2F.2, cut = .3, digits = 3, sort = TRUE)

reduced.model.2F.2 <- dplyr::select(efa.data ,
                                    -c( CS09, CS03, CS16, CS01, RCS05, RCS02, CS21, CS23))

correlations.red.2F.2 <- polychoric(reduced.model.2F.2)

fa.2F.3 <- fa(r=correlations.red.2F.2$rho, nfactors = 2, fm= "pa",rotate ="promax",
              residuals = TRUE, SMC = TRUE, max.iter = 500 )

CC <- print(fa.2F.3, cut = .3, digits = 3, sort = TRUE)

var1 <- CC$Vaccounted[2,1]*100
var2 <- CC$Vaccounted[2,2]*100

omega <- psych::omega(reduced.model.2F.2,2)

efa.table.1 <- tabledown::fac.tab(fa.2F.3,cut=.3, complexity = F)

#Communality
Fac.2 <- fa.2F.3
min.com <- min(Fac.2$communality)
max.com <- max(Fac.3$communality)

# loading
min.loadings <- (min(Fac.2$loadings))
max.loadings <- (max(Fac.2$loadings))

# Residuals
residual.5F =residuals(fa.2F.3, diag = FALSE, na.rm =T)
#Count of numbers of residuals>.05
BigRR= sum(residual.5F>abs(.05), na.rm =T)
print(BigRR)

#Total number of off diagonal elements in the data matrix
totRR = length(reduced.model.2F.2)*(length(reduced.model.2F.2)-1)/2

# Proportion of off-diagonal elements >.05
sumR <- sum(BigRR/totRR*100)

```

```{r tau-equivalance, eval=FALSE, include=FALSE}
tau.equivalance <- coefficientalpha::tau.test(reduced.model.2F.2)
```

```{r ordinal-alpha, include =T}
F1 <- dplyr::select(efa.data, CS15, CS22, CS20, CS11, CS14, CS18, CS13, CS06, CS10)
F2 <- dplyr::select(efa.data, CS19, CS04, CS17, CS12, CS08, CS07)

#ordinal alpha
F1.alpha <- psych::alpha(polychoric(F1)$rho,check.keys=TRUE)
F2.alpha <- psych::alpha(polychoric(F2)$rho,check.keys=TRUE)

F1.alpha <- F1.alpha$total$raw_alpha
F2.alpha <- F2.alpha$total$raw_alpha

```

```{r EFA2f, include=FALSE}
fa.1F.1 <- fa(r=correlations$rho, nfactors = 1, fm= "pa",
              residuals = TRUE, SMC = TRUE)
DD <- print(fa.1F.1, cut = .3, digits = 3, sort = TRUE)

reduced.model.1F.1 <- dplyr::select(efa.data ,
                                    -c( CS16, RCS02, RCS05, CS01, CS09, CS23))

correlations.red.1F.1 <- polychoric(reduced.model.1F.1)

fa.1F.2 <- fa(r=correlations.red.1F.1$rho, nfactors = 1, fm= "pa",
              residuals = TRUE, SMC = TRUE)

EE <- print(fa.1F.2, cut = .3, digits = 3, sort = TRUE)
var3 <- EE$Vaccounted[2,1]*100

one.factor.alpha <- psych::alpha(correlations.red.1F.1$rho ,check.keys=TRUE)
one.factor.alpha <- one.factor.alpha$total$raw_alpha

```

At first we fitted one factor solution in our dataset with all 23 items. The initial solution had six items with poor factor loadings (Item 1, 2,5,9, 16, 23). We discarded these items from the model and run another EFA. This iteration of EFA yielded a one factor simple structure with 17 items. The one factor solution explained `r apa(var3,2,T)`% of the total variance. Cronbach's alpha coefficient has a tendency to deflate the estimates for  Likert type data as the calculation is based on pearson-correlation matrix which requires data with continuous measurement level [@gadermann2012estimating; @zumbo2007ordinal]. For better estimates of reliability we reported ordinal alpha using polychoric-correlation [@zumbo2007ordinal]. The internal consistency reliability coefficient for this model was, ordinal $\alpha$ = `r apa(one.factor.alpha,2,F)`.

The initial two-factor solution with all 23 items had several items (item 01, 03, 05 09, 16) with poor factor loading (<.3) . We conducted another two round of EFA and identified and discarded problematic items (with poor factor loadings and/or cross loadings). Finally, a two-factor EFA solution with 15 items was accepted with RMSR = 0.06, no loading smaller than .30 and no cross-loading greater than .30. The first factor retained 8 items, and the second factor retained 6 items. The first factor explained  `r apa(var1,2,T)`% of the total variance and the second factor explained only `r apa(var2,2,T)`.  Ordinal alpha coefficient value ranges from 0 to 1 and higher value represents better reliability.  The internal consistency reliability coefficient for the two factors were, ordinal $\alpha$ = `r apa(F1.alpha,2,F)` and `r apa(F2.alpha,2,F)` respectively. Items clustered under factor 1 were stemmed from several common theme including efficiency of verbal communication,   individual's efficiency in verbal communication. All items except item12 (I try to see the other person's point of view) investigated individual's efficiency in non-verbal communication. The average variance extracted (AVE) is the amount of joint variance captured by the components and not by measurement error. @fornell1981evaluating recommended .50 as a benchmark value of AVE to establish the converging validity and purity of particular factor. AVE for  these two factors were.25 and .23 indicating high measurement error. Due to such high measurement error of the factor we rejected the two factor model and accepted the one factor model which had satisfactory reliability

```{r}
apa_table(efa.table.1, caption = "Factor loadings of two factor model")
```

```{r EFA2f, include=FALSE}
fa.1F.1 <- fa(r=correlations$rho, nfactors = 1, fm= "pa",
              residuals = TRUE, SMC = TRUE)
AA <- print(fa.1F.1, cut = .3, digits = 3, sort = TRUE)

reduced.model.1F.1 <- dplyr::select(efa.data ,
                                    -c( CS16, RCS02, RCS05, CS01, CS09, CS23))

correlations.red.1F.1 <- polychoric(reduced.model.1F.1)

fa.1F.2 <- fa(r=correlations.red.1F.1$rho, nfactors = 1, fm= "pa",
              residuals = TRUE, SMC = TRUE)

BB <- print(fa.1F.2, cut = .3, digits = 3, sort = TRUE)

```

```{r eval=FALSE, include=FALSE}
model.0 <- "F1 =~ CS15 + CS22+CS20+CS11+CS14+CS18+CS13+CS06+CS10
            F2 =~ CS19+CS04+CS17+CS08+CS07"
fit.0 <- cfa(model.0, data = cfa.data, estimator = "WLSMV")

## Summary of Model 
cfa.sum.0 <- summary(fit.0, fit.measures =TRUE,standardized = TRUE,rsq =TRUE)

## Selected Fit measures 
fit.measures.0 <- fitmeasures (fit.1,c("gfi", "agfi", "nfi","rfi", 
                       "cfi","tli",
                       "rmsea", "rmsea.ci.lower", "rmsea.ci.upper","srmr"))
reliability0 <- semTools::reliability(fit.1, return.total = T)

```

```{r CFAitems, include=FALSE }
#This chunk holds codes for 1st CFA

model.1 <- "Communication =~ CS15 + CS21 +CS11 + CS20+ CS14+ CS19+ CS22+ CS12+ CS06+ CS07+ CS18+ CS08+        CS04+CS17+CS03+CS13+CS10"

fit.1 <- cfa(model.1, data = cfa.data, estimator = "WLSMV")

## Summary of Model 
cfa.sum.1 <- summary(fit.1, fit.measures =TRUE,standardized = TRUE,rsq =TRUE)

## Selected Fit measures 
fit.measures.1 <- fitmeasures (fit.1,c("gfi", "agfi", "nfi","rfi", 
                       "cfi","tli",
                       "rmsea", "rmsea.ci.lower", "rmsea.ci.upper","srmr"))

reliability1 <- semTools::reliability(fit.1)

```

```{r}
anova(fit.0,fit.1)
```

```{r mod1, eval=FALSE, include=FALSE}
#This chunk holds codes for modification indices for 1st CFA
modfit.Cor.one <- modindices(fit.Cor.1, sort. = TRUE) 
modfit.Cor.one[modfit.Cor.one$mi>3.84,]
```

```{r cfaplot, include =F}
#This chunk holds codes to store CFA plot

png(filename="figures/CFAplot.png", 
    type="cairo",
    units="in", 
    width=6, 
    height=6, 
    pointsize=12, 
    res=1200)
semPaths (fit.1 , 
          what= "std",
          #"hide", #(hides coefficeits)
          #whatLabels = "std",
          intercepts = F,
          style ="OpenMx",
          #residScale = 6,
          theme = "colorblind",
          nCharNodes = 0,
          reorder =T,
          rotation =2,
          layout ="tree",
          cardinal = T,
          curvePivot =T,
          sizeMan =8,#items length
          sizeMan2 = 2,#items
          sizeLat = 12,#factors
          thresholds = FALSE,
          equalizeManifests =F,
          fade = FALSE,
          edge.label.cex = .8,
          #exoCov = T,
          centerLevels = T,
          #edge.color="black",
          label.scale=T,
          label.cex=1.2, #Font of factor and item name
          residuals=FALSE,
           #exoVar=FALSE,
          #fixedStyle=6, #Style of arrow (guide item)
          #freeStyle=1#Style of arrow (other item), #XKCD = TRUE

          )

dev.off()
```

```{r Shortcut Invariance, eval=FALSE, include=FALSE}
invariance.data <- cfa.data.full[, c(3, 6:28)]
invariance.data <- na.omit(invariance.data )
library("semTools")
invariance.model <- measurementInvariance(model = model.1,
                                          data = invariance.data,
                                          group = "Gender",                                                                         estimator = "WLSMV",
                                          strict = T)
invariance.model[[1]] #configural Invariance
invariance.model[[2]] # loading, weak, metric
invariance.model[[3]] # intercept, strong, scaler
invariance.model[[4]] # residual, strict
invariance.model[[5]] # means, structural invariance
```

```{r figcfa, fig.cap = "CFA Plot.", warning=FALSE}

knitr::include_graphics("Figures/CFAplot.png", dpi =600)

```

## Measurement Invariance
```{r InvarianceAnalysis_data, include=FALSE}
#This chunk holds codes to prepare measurement invariance data.

invariance.data <- cfa.data.full[, c(3, 6:28)]
invariance.data <- na.omit(invariance.data )
mi.count <- invariance.data %>% 
  group_by(Gender) %>% 
  count() %>% 
  as.data.frame()
```

```{r Invariancedetail, include=FALSE}
#This chunk holds codes for measurement Invariance
library(lavaan)
model.1 <- "Communication =~ CS15 + CS21 +CS11 + CS20+ CS14+ CS19+ CS22+ CS12+ CS06+ CS07+ CS18+ CS08+        CS04+CS17+CS03+CS13+CS10"

# Configural invariance ####
configural <- cfa(model = model.1,
                  data = invariance.data,
                  group = "Gender",
                  estimator = "WLSMV")

summary(configural, fit.measures =TRUE,standardized = TRUE, rsq =TRUE)

fitmeasures (configural,c("gfi", "agfi", "nfi","rfi", 
                       "cfi","tli",
                       "rmsea", "rmsea.ci.lower", "rmsea.ci.upper","srmr"))

#Metric
weak <- cfa(model = model.1,
                  data = invariance.data,
                  group = "Gender",

                  estimator = "WLSMV", 
            group.equal = "loadings")

summary(weak, fit.measures =TRUE,standardized = TRUE, rsq =TRUE)

fitmeasures (weak,c("gfi", "agfi", "nfi","rfi", 
                       "cfi.scaled","tli.scaled",
                       "rmsea.scaled", "rmsea.ci.lower.scaled", "rmsea.ci.upper.scaled","srmr"))

First.comp <- compareFit (configural, weak)
lavaan::summary(First.comp)

#Scaler
strong <- cfa(model = model.1,
                  data = invariance.data,
                  group = "Gender",

                  estimator = "WLSMV", 
              group.equal = c("loadings", "intercepts"))
summary(strong, fit.measures =TRUE,standardized = TRUE, rsq =TRUE)

Second.comp <- compareFit  (weak, strong)
summary(Second.comp)

#residual
strict <- cfa(model = model.1,
                  data = invariance.data,
                  group = "Gender",

                  estimator = "WLSMV",  
              group.equal = c("loadings", "intercepts", "residuals"))

summary(strict, fit.measures =TRUE,standardized = TRUE, rsq =TRUE)
residual.fitm <- as.data.frame(fitmeasures (strict,c("gfi", "agfi", "nfi","rfi", "cfi","tli","rmsea", "srmr","aic")))

Third.comp <- compareFit(strict,strong)
summary(Third.comp)
#structural
structural <-cfa(model = model.1,
                  data = invariance.data,
                  group = "Gender",

                  estimator = "WLSMV",  
                 group.equal = c("loadings", "intercepts", "residuals", "means","lv.variances","lv.covariances"))

summary(structural, fit.measures =TRUE,standardized = TRUE, rsq =TRUE)
fitmeasures (structural,c("gfi", "agfi", "nfi","rfi", "cfi","tli","rmsea", "srmr","aic"))

fourth.comp <- compareFit  (structural,strict)
summary(fourth.comp)
comfit.par <- compareFit(configural, weak, strong, strict)
summary (comfit.par)

models <-  list("Configural" = configural, 
                "Metric" = weak, 
                "Scalar" = strong, 
                "Residual" = strict)

Invariance.table <- compareLavaan(models,
              nesting = "Configural > Metric > Scalar > Residual", 
              fitmeas = c("chisq", "df",  "cfi","tli","rmsea", "rmsea.ci.lower", "rmsea.ci.upper", "srmr" ),
              scaled = F,
              chidif = T, digits = 2)

colnames(Invariance.table) <- c("Chi-Square", "df", "CFI", "TLI", "RMSEA", "RMSEA 90% Lower CI", "RMSEA 90% Upper", "Chi-Square Difference", "df difference*", "p")

```

```{r InvarianceTab, results='asis'}
apa_table(Invariance.table, align = "c",  caption = "Measurment Invariance analysis on CFA sample (n=262) across native and non-native English speakers.", note = " a = Metric vs Configural; b = Scalar vs Metric; c = Residual vs Scalar; * =  df of model comparison", landscape = T, font_size = "footnotesize" )

```

```{r samplesizeIRT, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE, results='hide', width=60}
set.seed(222)
m <-  23 # item number
n <- c(50, 75, 100, 150, 200, 300, 350) # sample size
design <- as.data.frame(n)

irtGenerate <- function(condition, fixed_objects = FALSE) {
n <- condition$n

a <- matrix(rlnorm(m,.2,.3))
diffs <- t(apply(matrix(runif(m*4, .3, 1), m), 1, cumsum))
diffs <- -(diffs - rowMeans(diffs))
diffs<- diffs + rnorm(m)
d <- diffs
dat <- simdata(a, d, n, itemtype = 'graded')
return(dat)
}

irtAnalyze <- function(condition, dat, fixed_objects = NULL) {
mod <- mirt(dat, 1, itemtype = 'graded', verbose = FALSE)
simpars <- coef(mod, simplify = TRUE, digits = Inf)$items
irtpars <- c(a = simpars[,1], d = simpars[,2])
return(irtpars)
}

irtSummarize <- function(condition, results,
fixed_objects = NULL) {
apop <- fixed_objects['alpha', ]
dpop <- fixed_objects['d', ]
simrmse <- RMSE(results, c(apop, dpop))
out <- c(RMSE = simrmse)
return(out)
}

simres <- runSimulation(design, replications = 100,
parallel = TRUE, generate = irtGenerate,
analyse = irtAnalyze, summarise = irtSummarize,
packages = c('mirt'))
simres

colind <- grep(".a.", colnames(simres))
sima <- as.data.frame(simres[, colind])
nvec <- as.numeric(simres$n)

meanRMSE <- rowMeans(sima)
names(meanRMSE) <- n
round(meanRMSE, 2)
matplot(nvec, log(sima), type = "l", col = 1, lty = 1,
ylab = "log(RMSE)", xlab = "sample size",
main = "Graded Response Model", xaxt = "n")
axis(1, at = nvec)
```

To gather more information on our retained one-factor solution, we sought Item Response Theory (IRT). IRT complements the conventional classical test theory-based analysis by gathering information on item discrimination and item difficulty. IRT judges an item's quality by providing item information in the light of participants' trait level ($\theta$). We gathered evidence on item quality as well as item fit, person fit and model by fitting a graded response model  in RStudio with the "mirt" package [@R-mirt]  [@R-mirt]. We did a Monte Carlo simulation using "SimDesign" package [@R-SimDesign]  with sample sizes varying from 50-350 and calculated average root mean squared error(RMSE) to estimate the optimal sample size for the graded response model with 23 items. The RMSE became stable for n = 200 to 300 (RMSE ranging between .25-.35). Our sample size within the estimated sample size for stability.

```{r, include =F}
irt.data <- na.omit(irt.data)
irt.data.1 <- irt.data %>% 
  dplyr::select( CS03, CS04, CS06, CS07,CS08,CS10,CS11, CS12, 
                 CS13,CS14,CS15,CS17, CS18, CS19, CS20 ,CS21, CS22)

fit <- mirt(irt.data.1, model = 1, itemtype = 'graded', 
               SE = TRUE, Se.type = 'MHRM',
               technical = list(NCYCLES = 10000))

# Model Parameters ####
params <- coef(fit, IRTpars = TRUE, simplify = TRUE, rawug = FALSE) 
items <- data.frame(params$items)
se <- coef(fit, printSE = TRUE)

# Model Fit (degrees of freedom too low to check model fit)
model <- M2(fit)

#Item fit
item.fit<- itemfit(fit, fit_stats = c("S_X2", "G2","Zh", "infit"),
                                  impute=10)

#Identifying Missfit items based on RMSEA Value
item_misfits <- subset(item.fit, RMSEA.S_X2 >= .06)

## Creating IRT table
item.fit2 <- item.fit[,c(4,6,11,12,13,14)] 
colnames(item.fit2) <- c("Standardized Outfit", "Standardized Infit", "S-Chi-square", "df", "RMSEA", "p")

IRT.details <- cbind(items ,item.fit2)

IRT.details <- IRT.details %>% 
 dplyr::mutate_if(is.numeric, round, digits=3)

IRT.details <- tibble::rownames_to_column(IRT.details, "Items")
#write_csv(IRT.details, "table_raw/IRT_parameters.csv")

#-3<b<3
#.5<a<2; baker p12

mean.discrimination <- mean(IRT.details$a)
sd.discrimination <- sd(IRT.details$a)
range.descrimination <- range(IRT.details$a)

cat.prob <- IRT.details[,c(3:6)]
cat.prob.long <- reshape::melt(cat.prob)
range.difficulty <- range(cat.prob.long$value)

range.outfit <-  range(IRT.details$Outfit)
range.infit <-  range(IRT.details$Infit)
range.rmsea <- range(IRT.details$RMSEA)

# Person Fit ####

personfit <- personfit(fit) 
personfit_model_misfits <- subset(personfit, Zh < -2)
rownames(personfit_model_misfits)
nrow(personfit_model_misfits)
hist(personfit$Zh, xlab = "Zh Statistics",main = "F1 Person Fit")
abline(v = -2, lwd = 2, lty = 2)

person <- ggplot(personfit, aes(x=Zh)) + geom_histogram(binwidth=.5,col=I("black"),fill="palegreen4")+geom_vline(xintercept = -2)+
  labs( y="Number of Participants", x = "Zh Statistics")+apatheme + xlim(-4,2)+
   scale_y_continuous(limits = c(0, 40), expand = expand_scale(mult = c(0, 0.1)))

# annotation.1 <- data.frame(
#    x = (-3.1),
#    y = (20),
#    label = ("Most of the participants
#         are within the guidline")
# )
# 
# per.anno <-person + geom_label(data=annotation.1, aes( x=x, y=y, label=label),                 , 
#            color="orange", 
#            size=5 , angle=45, fontface="bold" )

ggsave("Figures/person.png",person , width = 12, height = 8, dpi = 600, bg = "white")

# Plots ####

OCC <- plot(fit, type = "trace", facet =T, main = "Communication Scale OCC")
info <- plot(fit, type = "infoSE", main = "Communication Scale")
iteminfo <- plot(fit, type = "infotrace", 
                    facet_items = T, main = "Communication Scale")

#conditional reliability
reliability <- plot(fit, type = 'rxx', theta_lim = c(-6, 6), 
      main="" )

marginal.rel <- marginal_rxx(fit)

#scale characteristics curve
scale <- plot(fit, type = 'score', theta_lim = c(-6, 6), main = "")

#-3<b<3
#.5<a<2; baker p12
```

```{r irttab, results='asis'}
papaja::apa_table(IRT.details, caption = "Items discrimination and response category difficulty thresholds of 17 items in  Bangla Communication Scale (n =581)", note = "a = item discrimination parameter; b(1-4) = response category difficulty parameter")
```

```{r OCC, include=F}
ggicc <- function(model, item, theta){
  Theta <- matrix(seq(-theta,theta, by = .1))
  iteminfo <- mirt::extract.item(model, item)
  P <-mirt::probtrace(x=iteminfo , Theta=Theta  )
  icc <- data.frame(P = P, Theta = Theta)
  colnames(icc) <- c(paste("P", 1:ncol(P), sep=''), "Theta")
  icc2<- reshape(icc, direction='long', varying = paste("P", 1:ncol(P), sep=''), v.names = 'P',
                 times = paste("P", 1:ncol(P), sep=''))
  plot <- ggplot2::ggplot(icc2,aes(Theta,P, col =time))+geom_line()+xlab(expression(theta)) + 
    ylab(expression(P(theta)))+ theme(legend.title=element_blank())
  return(plot)

}

library(ggsci)

item03 <- ggicc(fit,1,6)+apatheme+ggtitle("item03")+apatheme+scale_color_npg()+
  scale_y_continuous(breaks=c(0,.5,1), limits = c(0, 1))

item04 <- ggicc(fit,2,6)+apatheme+ggtitle("item04")+apatheme+scale_color_npg()+
  scale_y_continuous(breaks=c(0,.5,1), limits = c(0, 1))

item06 <- ggicc(fit,3,6)+apatheme+ggtitle("item06")+apatheme+scale_color_npg()+
  scale_y_continuous(breaks=c(0,.5,1), limits = c(0, 1))

item07 <- ggicc(fit,4,6)+apatheme+ggtitle("item07")+apatheme+scale_color_npg()+
  scale_y_continuous(breaks=c(0,.5,1), limits = c(0, 1))

item08 <- ggicc(fit,5,6)+apatheme+ggtitle("item08")+apatheme+scale_color_npg()+
  scale_y_continuous(breaks=c(0,.5,1), limits = c(0, 1))

item10 <- ggicc(fit,6,6)+apatheme+ggtitle("item10")+apatheme+scale_color_npg()+
  scale_y_continuous(breaks=c(0,.5,1), limits = c(0, 1))

item11 <- ggicc(fit,7,6)+apatheme+ggtitle("item11")+apatheme+scale_color_npg()+
  scale_y_continuous(breaks=c(0,.5,1), limits = c(0, 1))
ggsave("Figures/ideal_icc.png",item11 , width = 12, height = 8, dpi = 600, bg = "white")

item12 <- ggicc(fit,8,6)+apatheme+ggtitle("item12")+apatheme+scale_color_npg()+
  scale_y_continuous(breaks=c(0,.5,1), limits = c(0, 1))

item13 <- ggicc(fit,9,6)+apatheme+ggtitle("item013")+apatheme+scale_color_npg()+
  scale_y_continuous(breaks=c(0,.5,1), limits = c(0, 1))

item14 <- ggicc(fit,10,6)+apatheme+ggtitle("item14")+apatheme+scale_color_npg()+
  scale_y_continuous(breaks=c(0,.5,1), limits = c(0, 1))

item15 <- ggicc(fit,11,6)+apatheme+ggtitle("item15")+apatheme+scale_color_npg()+
  scale_y_continuous(breaks=c(0,.5,1), limits = c(0, 1))

item17 <- ggicc(fit,12,6)+apatheme+ggtitle("item17")+apatheme+scale_color_npg()+
  scale_y_continuous(breaks=c(0,.5,1), limits = c(0, 1))

item18 <- ggicc(fit,13,6)+apatheme+ggtitle("item18")+apatheme+scale_color_npg()+
  scale_y_continuous(breaks=c(0,.5,1), limits = c(0, 1))

item19 <- ggicc(fit,14,6)+apatheme+ggtitle("item19")+apatheme+scale_color_npg()+
  scale_y_continuous(breaks=c(0,.5,1), limits = c(0, 1))

item20 <- ggicc(fit,15,6)+apatheme+ggtitle("item20")+apatheme+scale_color_npg()+
  scale_y_continuous(breaks=c(0,.5,1), limits = c(0, 1))

item21 <- ggicc(fit,16,6)+apatheme+ggtitle("item21")+apatheme+scale_color_npg()+
  scale_y_continuous(breaks=c(0,.5,1), limits = c(0, 1))

item22 <- ggicc(fit,17,6)+apatheme+ggtitle("item22")+apatheme+scale_color_npg()+
  scale_y_continuous(breaks=c(0,.5,1), limits = c(0, 1))

ICC <- cowplot::plot_grid(item03, item04, item06, item07, item08, 
                   item10, item11, item12, item13, item14, item15,  item17, item18, item19,
                   item20, item21, item22,
                      labels = "AUTO",
                     ncol=3,  
                    label_size = 15, 
                     align = "v")

ggsave("Figures/ICC.png",ICC , width = 12, height =14, dpi = 600, bg = "white")

```

```{r figICC, eval=FALSE, fig.cap="Option Charecteritics Curve.", warning=FALSE, include=FALSE, out.height='100%'}

knitr::include_graphics("Figures/ICC.png", dpi =600)

```

```{r Test-info, include =F}

# iteminfo <- plot(fit, type = "infotrace",which.item =2)
# 
# itemplot.01 <- tabledown::ggiteminfo(fit,1,6)+apatheme+
#   ggtitle(" Item03")+theme(legend.position = "none")+
#   labs(y = "Item Information")+
#   scale_color_npg()+geom_hline(yintercept = .20, colour ="red")+geom_area(fill ="palegreen4" )+
#   xlab(expression(theta))+ ylab(expression(I(theta)))
# 
# 
# itemplot.02 <- tabledown::ggiteminfo(fit,2,6)+apatheme+
#   ggtitle(" Item04")+theme(legend.position = "none")+
#   labs(y = "Item Information")+
#   scale_color_npg()+geom_hline(yintercept = .20, colour ="red")+geom_area(fill ="palegreen4" )+
#   xlab(expression(theta))+ ylab(expression(I(theta)))
# 
# 
#

# good.info <- cowplot::plot_grid( item6.iic1, item7.iic1, item11.iic1, item12.iic1, item14.iic1, item15.iic1 , item18.iic1, item19.iic1, item20.iic1, item21.iic1, item22.iic1, 
#                                  labels = "AUTO",ncol=3, label_size = 15,align = "v")
# 
# 
# ggsave("Figures/good.info.png",good.info, width = 12, height = 8, dpi = 600, bg = "white")

test.info <- tabledown::ggtestinfo(irt.data.1, fit)+ apatheme+geom_line(colour = "black", size=1.5)+geom_area(fill ="#E64B3599" )
# +geom_vline(xintercept = -3, colour ="black")+
#   geom_vline(xintercept = .5, colour ="black")
ggsave("Figures/TIC.png",test.info , width = 6, height = 4, dpi = 600, bg = "white")

library(plotly)
info.p <- plotly::ggplotly(test.info)

```

```{r figTIC, fig.cap = "TIC.",  out.height='100%', warning=FALSE}

knitr::include_graphics("Figures/TIC.png", dpi =600)

```

```{r}
# This chinks holds code of compiled item info code created using facet wrap. (Not included in manuscript)
pacman::p_load(reshape,ggsci)
Theta <- matrix(seq(-6,6, by = .1))

##Item Extraction
item03 <- extract.item(fit, 1)#item number
item04 <- extract.item(fit, 2)
item06 <- extract.item(fit, 3)
item07 <- extract.item(fit, 4)
item08 <- extract.item(fit, 5)
item10 <- extract.item(fit,6)
item11<- extract.item(fit, 7)
item12 <-extract.item(fit, 8)
item13 <- extract.item(fit, 9)
item14 <- extract.item(fit, 10)
item15 <- extract.item(fit, 11)
item17 <- extract.item(fit, 12)
item18 <- extract.item(fit, 13)
item19 <- extract.item(fit, 14)
item20 <- extract.item(fit, 15)
item21 <- extract.item(fit, 16)
item22 <- extract.item(fit, 17)

##Item Info
Item03 <- iteminfo(item03, Theta)
Item04 <- iteminfo(item04, Theta)
Item06 <- iteminfo(item06, Theta)
Item07 <- iteminfo(item07, Theta)
Item08 <- iteminfo(item08, Theta)
Item10 <- iteminfo(item10, Theta)
Item11 <- iteminfo(item11, Theta)
Item12 <- iteminfo(item12, Theta)
Item13 <- iteminfo(item13, Theta)
Item14 <- iteminfo(item14, Theta)
Item15 <- iteminfo(item15, Theta)
Item17 <- iteminfo(item17, Theta)
Item18<- iteminfo(item18, Theta)
Item19 <- iteminfo(item19, Theta)
Item20 <- iteminfo(item20, Theta)
Item21 <- iteminfo(item21, Theta)
Item22 <- iteminfo(item22, Theta)

#Data frame

item.info.data <- as.data.frame(cbind(Item03, Item04,Item06, Item07, Item08,
                                      Item10, Item11, Item12, Item13, Item14,
                                      Item15, Item17, Item18, Item19, Item20,
                                      Item21, Item22))

item.info.data.long <- melt(item.info.data)

item.info.data.2 <- data.frame(Theta=seq(-6,6, by = .1), item.info.data.long)

pacman::p_load(ggsci)

item.info.plot <- ggplot(item.info.data.2,aes(Theta,value,col=variable))+geom_line()+apatheme+
  ggtitle("Item Information Plot")+theme(legend.position = "none")+
  facet_wrap(~variable)+theme(strip.text.x = element_text(size = 15))+geom_area()+ scale_color_npg()+aes(fill = as.factor(variable))+xlab(expression(theta)) + 
  ylab(expression(I(theta)))

# facet_wrap(~variable,scales = "free")++geom_hline(yintercept = .20, colour ="red")

ggsave("Figures/itemInfoplot.png",item.info.plot, width = 8, height =8, dpi = 1200, bg ="white")

```

```{r figIIC, fig.cap = "Item Information Curve.",  out.height='100%',out.width='100%', warning=FALSE}

knitr::include_graphics("Figures/itemInfoplot.png", dpi =600)

```

```{r conrel-print, include=F}
##
png(filename="Figures/conrel.png", 
    type="cairo",
    units="in", 
    width=12, 
    height=2, 
    pointsize=12, 
    res=300)
library(latticeExtra)
library(directlabels)
font.settings <- list(
  font = 1,
  cex = 6,
  fontfamily = "sans")

apa.latice  <- list(
  par.xlab.text = font.settings,
  par.ylab.text = font.settings,
  axis.text = font.settings,
  sub.text = font.settings,
  add.text = font.settings)

 plot(fit, type = 'rxx', theta_lim = c(-6, 6), 
      main="Conditional Reliability Plot", par.settings=c(apa.latice),col.line = "#E69F00",lwd = 3.5 )

dev.off()
```

```{r figconrel, fig.cap = "Conditional Reliability.", out.width= '100%', out.height='50%', warning=FALSE}

knitr::include_graphics("Figures/conrel.png", dpi =600)

```

Marginal reliability is based on the true score model (Lord & Novick, 1968) and is an estimate of the overall reliability of a test based on the average conditional standard errors. Often it is close in value to coefficient alpha (and sometimes it may even be identical). Alpha provides a lower estimate of marginal reliability.

```{r dif, eval=FALSE, include=FALSE}
library("lordif")
gender <- as.data.frame(data$Gender)
colnames(gender) <- "Gender"
gender$Gender <- as.factor(gender$Gender)
irt.data <- na.omit(irt.data)
communication <- sem.data[,c(2:24)]
dif.data <- cbind(gender, communication)
dif.data <- na.omit(dif.data)
dif.data.full <- as.data.frame(dif.data)
DIF <- lordif(as.data.frame(irt.data), group = dif.data.full$Gender, criterion = "Chisqr")

DIF$stats[,1:5]
plot(DIF)
```

```{r}

val.data<- validity.data.full %>% 
  dplyr::select( CS03, CS04, CS06, CS07,CS08,CS10,CS11,    CS12, 
                 CS13,CS14,CS15,CS17, CS18, CS19, CS20 ,CS21, CS22) %>% 
  rowwise() %>% 
  mutate(Total_communication = sum(across(starts_with("CS")), na.rm = T),
         F1 = sum(CS15,CS22,CS20,CS14,CS18,CS13,CS06,CS10, CS11),
         F2 = sum(CS19,CS04,CS17,CS08,CS07))

val.data.2 <- as.data.frame(cbind(val.data$Total_communication,val.data$F1, val.data$F2, validity.data.full$Total_HS, validity.data$Total_LS, validity.data.full$Total_SE))

colnames(val.data.2) <- c("Communication","F1", "F2", "Hopelessness", "Life Satisfaction", "SE")
validity <- na.omit(val.data.2)
```

```{r correlationfunc, include =F}
glrstab<- function(x, export=FALSE) {

 r <-corr.test(x)$r	#taking just the correlation matrix; no N, or p
 p <-corr.test(x)$p	#taking the p*s

#define notions for significance levels
 mystars <- ifelse(p < .001, "**"
                   , ifelse(p < .01, "**"
                            , ifelse(p < .05, "*"
                                     , ifelse(p < .10, "+", " "))))

 #round r, define new matrix Rnew with the correlations from rnd and paste mystars
 rnd  <- papaja::printnum(r, gt1 = FALSE, digits = 2)  #round, drop leading 0 - Thanks CRSH!								                     
 Rnew <- matrix(paste(rnd, mystars, sep=""), ncol=ncol(rnd))

  #remove 1.0 correlations from diagonal  and set the strings
 diag(Rnew) <- ''		
 Rnew[upper.tri(Rnew)] <- ''

 rownames(Rnew) <- paste(1:ncol(rnd), colnames(rnd), sep=" ")         #define number and name
 colnames(Rnew) <- paste(1:ncol(rnd), "", sep="") 			       #define number

#fun-part: we trim the top half 
 Rnew[upper.tri(Rnew)] <- ''			
 Rnew

 #Rnew <- cbind(round(describe(x)[,3:4],2), Rnew)		     #describe x, M sD - put them in the matrix
 #colnames(Rnew)[1:2] <- c("M","SD")					      		#Beschriftung der neuen Spalten
 Rnew <- Rnew[,1:(ncol(Rnew)-1)]							        	#delete the last column (ugly)

 #export to clipboard

   if (export==TRUE){
   result<-write.table(Rnew
                       , "clipboard"
                       , sep=";"
                       , row.names=FALSE)
 }
 else result <- Rnew
 return(result)

}

```

```{r correlations, include =F}
Cor <-glrstab(validity) #the function in action!

cor.tab <- as_tibble(cor(validity))

```

```{r tabvalidity,  results='asis'}
papaja::apa_table(Cor, escape  = FALSE
                , caption = "Correlation matrix of the main variables"
                , note    = "**p < .001")
```

# Results

# Discussion

## Ethical Consideration

All procedures performed in studies involving human participants were in accordance with the 1964 Helsinki declaration and its later amendments or comparable ethical standards. This article does not contain any studies with animals performed by any of the authors.

## Data and code availability
All code and data underlying this article is available on a public GitHub repository (https://github.com/masiraji/Communication).

\newpage

# References

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id="refs" custom-style="Bibliography"></div>
\endgroup